### 2011-08-19 src/main/latex/ebm/ebm.tex
Porting to mercurial

\documentclass{scrartcl}

\usepackage[british]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}  %theorem definition dependency
%\usepackage{moreverb} %source code listing dependency
%\usepackage{textcomp} %source code listing dependency
%\usepackage{listings} %source code listing dependency
\usepackage{color}    %source code listing dependency
\usepackage{titling}  %pdf metadata dependency
\usepackage{ifpdf}    %pdf metadata dependency
%\usepackage{graphicx}
%\usepackage{caption}
%\usepackage{pstricks}
\usepackage{forloop}

%% DOCUMENT METADATA %%

\title{Integrating experimental networks in the absence of Gold Standards}
\author{Jochen Weile, M.Sc.}
\date{\today}

%% LAYOUT OPTIONS %%
\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

%% CONFIGURE CODE HIGHLIGHTING %%
%\lstset{
%	tabsize=4,
%	language=Java,
%  basicstyle=\footnotesize,
%  upquote=true,
%  aboveskip={1.5\baselineskip},
%  columns=fixed,
%  showstringspaces=false,
%  extendedchars=true,
%  breaklines=true,
%  prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
%  frame=none,
%  showtabs=false,
%  showspaces=false,
%  showstringspaces=false,
%  identifierstyle=\ttfamily,
%  keywordstyle=\bf\color[rgb]{0.5,0,0.7},
%  commentstyle=\color[rgb]{0.5,0.5,0.5},
%  stringstyle=\color[rgb]{0,0,1},
%}

%% SUPPORT FOR PDF METADATA %%
\ifpdf
	\pdfinfo {
		/Author (\theauthor)
		/Title (\thetitle)
		/CreationDate (D:20110216120000)
	}
\fi

%% DEFINE MACROS %%
\newcommand{\prob}{\mathbb{P}}
\newcommand{\odds}{\mathbb{O}}
\newcommand{\likeli}{\mathcal{L}}
\newcommand{\model}{\mathcal{H}}
\newcommand{\p}{\text{p}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\pardiff}[1]{\frac{\partial}{\partial #1}}
\newcommand{\remark}[1]{\begin{center}\fcolorbox{black}{yellow}{\parbox{.8\textwidth}{#1}}\end{center}}

%% DEFINE ENVIRONMENTS %%
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\section{Introduction}

\paragraph{Aim:} We are given $n$ experimentally measured networks, which we want to integrate. We want to assign an existence probability to each edge in our integrated network, based on the qualities of the experimental networks.

\section{Basic notions and definitions}

%\subsection{Graphs}

%\begin{definition}
%An (undirected) graph or network is a pair $G=(V,E)$ where $V$ is the set of vertices (nodes) and $E \subseteq \binom{V}{2}$ is the set of edges(arcs).
%\end{definition}

%\subsection{Probabilistics}

%\begin{definition}[Conditional probability]
%Let $A$ and $B$ be two events, then $$\prob(A|B)$$ denotes the conditional probability of event $A$ given the occurrence of event $B$. 
%\end{definition}

%\begin{theorem}[Bayes's theorem]
%Let $A$ and $B$ be two events, then $$\prob(A|B)=\frac{\prob(B|A)\cdot \prob(A)}{\prob(B)}$$
%\begin{description}
% \item[$\prob(A|B)$] is called the posterior probability of $A$ given $B$; 
% \item[$\prob(B|A)$] is called the likelihood of $B$ given $A$; 
% \item[$\prob(A)$] is called the prior probability of $A$; and 
% \item[$\prob(B)$] is called the marginal probability of $B$.
%\end{description}
%\end{theorem}

%\begin{definition}[Odds]
%The odds of an event $A$ are defined as $$ \odds(A) = \frac{\prob(A)}{\prob(\overline{A})} $$
%\end{definition}

\subsection{Experimental networks}
Let 
$$G = (V, E_G) \text{ , with vertices } V \text{ and edges } E_G \subseteq \binom{V}{2}$$ 
be a true network. 
Furthermore, let $X = (X_1, \ldots , X_n)$ be a vector of $n$ networks, that have been experimentally derived from $G$, where
$$X_i = (V, E_{X_i}) \text{ with the same vertices } V \text{ and different edges } E_{X_i} \subseteq \binom{V}{2} \,\,\forall\, 1 \le i \le n$$

Now, consider a potential edge $e \in \binom{V}{2}$. Then,
\begin{description}
	\item[$e \in E_{X_i}$] means that the experiment $X_i$ predicts that the edge $e$ exists.
	\item[$e \notin E_{X_i}$] means that the experiment $X_i$ predicts that the edge $e$ \textbf{does not} exist.
	\item[$e \in E_G$] means that the edge $e$ really exists.
	\item[$e \notin E_G$] means that the edge $e$ \textbf{does not} really exist.
\end{description}

\subsection{Edge prediction as a random event}
We can model the prediction of potential edge $e \in \binom{V}{2}$ as a random event. Let $D_i$ be a random variable, that assumes realisation 1 when $X_i$ predicts $e$ to exist and assumes realisation 0 when $X_i$ predicts $e$ to not exist. Let $$d_i = \mathbb{I}(e \in E_{X_i})$$ be the measured realisation of $D_i$, then $(D_i = d_i)$ is the event that $D_i$ assumes our measured realisation $d_i$.
Finally, let $$L \equiv (e \in E_G)$$ be the event that the edge really exists.

\section{Methods}

\subsection{Inferring edge probability from singular evidence}

We want to define the probability for an edge to really exist based on the prediction $d_i$ our experiment made. That is, we are interested in $\prob(L|D_i = d_i)$. For simplicity's sake we will abbreviate $\prob(L|D_i = d_i)$ as $\prob(L|d_i)$ in the following. Bayes's theorem gives us:

$$\prob(L|d_i) = \frac{\prob(d_i|L)\cdot \prob(L)}{\prob(d_i)}$$

However, since this form of notation can be more complicated than necessary for what we want to do in the following, we are converting our probabilities into \textit{odds} and apply Bayes's theorem again.

\begin{align}
\odds(L|d_i) &= \frac{\prob(L|d_i)}{\prob(\overline{L}|d_i)} = \frac{\frac{\prob(d_i|L)\cdot \prob(L)}{\prob(d_i)}}{\frac{\prob(d_i|\overline{L})\cdot \prob(\overline{L})}{\prob(d_i)}} \notag\\
 &= \frac{\prob(d_i|L)\cdot \prob(L)}{\prob(d_i|\overline{L}) \cdot \prob(\overline{L})} \notag\\
 &= \frac{\prob(d_i|L)}{\prob(d_i|\overline{L})} \cdot \odds(L) 
 \label{deduction_odds_edge}
\end{align}

$\odds(L)$ is the prior odds of an edge to really exist. We can estimate this number, but we will talk about that later. More important is the Bayes factor $\frac{\prob(d_i|L)}{\prob(d_i|\overline{L})}$. Let us call it $\Lambda_i$:

\begin{equation}
\label{def_lambda}
\Lambda_i := \frac{\prob(d_i|L)}{\prob(d_i|\overline{L})}
\end{equation}

Applying that definition \eqref{def_lambda} to our above equation \eqref{deduction_odds_edge}, we can write it down like this:

\begin{equation}
\label{odds_of_edge_pos}
\odds(L|d_i) = \Lambda_i \cdot \odds(L)
\end{equation}

\subsubsection{The Bayes factor}
\label{bayes_factor}

We see that the Bayes factor $\Lambda_i$ plays an important role. What exactly is it? 
\begin{itemize}
	\item In the case of $d_i = 1$
	\begin{itemize}
		\item The numerator $\prob(d_i = 1|L)$ is the probability of $X_i$ correctly reporting the edge that really exists. That is, the probability of producing a True Positive.
		\item The denominator $\prob(d_i = 1|\overline{L})$ is the probability of $X_i$ incorrectly reporting the edge, although it does not really exist. That is, the probability of producing a False Positive.
	\end{itemize}
	\item In the case of $d_i = 0$
	\begin{itemize}
		\item The numerator $\prob(d_i = 0|L)$ is the probability of $X_i$ incorrectly not reporting the edge although it really exists. That is, the probability of producing a False Negative. 
		\item The denominator $\prob(d_i = 0|\overline{L})$ is the probability of $X_i$ correctly not reporting the edge that does not really exist. That is the probability of producing a True Negative.
	\end{itemize}
\end{itemize}

So, suppose we knew the following numbers:

\begin{description}
	\item[$\alpha_i$]: The false positives rate in experiment $i$. That is, the share of false positives in the set of all not really existing edges..
	\item[$\beta_i$]: The false negative rate in experiment $i$. That is, the share of false negatives in the set of all really existing edges.
\end{description}

Then we could estimate $\Lambda_i$ as follows:

$$
\Lambda_i = \frac{\prob(d_i|L)}{\prob(d_i|\overline{L})} \approx
\begin{cases}
	d_i=1 & \frac{1-\beta_i}{\alpha_i}\\
	\text{otherwise} & \frac{\beta}{1-\alpha}
\end{cases}
$$

\subsection{Inferring edge probability from multiple evidence}

Now, we are interested in the odds of our edge to really exist considering \textbf{all} the different experimental networks. To answer this question we look at the problem from a recursive perspective.

Let $$D^{(n)} = \bigwedge_{i=1}^n (D_i = d_i)$$ be the combined event of observing the existence or non-existence of edge $e$ in all of our experimental networks.

Then, the odds $\odds(L|D^{(n)})$ for an edge existing given the predictions from all $n$ experiments is the same as the edge existing given the $n$'th network $D_n$ \textbf{and} the set of the remaining ones $D^{(n-1)}$. Using \eqref{odds_of_edge_pos} We can deduce:

\begin{align*}
\odds(L | D^{(n)} ) &= \odds(L | d_n \wedge D^{(n-1)}) \\
 &= \frac{\prob( d_n \wedge D^{(n-1)} | L) \cdot \prob(L)}{\prob( d_n \wedge D^{(n-1)} | \overline{L}) \cdot \prob(\overline{L}) } \\
 &= \frac{\prob( d_n |L) \cdot \prob( D^{(n-1)} | L) \cdot \prob(L)}{\prob( d_n |L) \cdot \prob( D^{(n-1)} | \overline{L}) \cdot \prob(\overline{L}) } \\
 &= \Lambda_n \cdot \frac{\prob( D^{(n-1)} | L)}{\prob( D^{(n-1)} | \overline{L})} \cdot \odds(L)\\
 &= \Lambda_n \cdot \frac{\prob(L| D^{(n-1)}) \cdot \prob(D^{(n-1)}) \cdot \prob(\overline{L})}{\prob(\overline{L} | D^{(n-1)}) \cdot \prob(D^{(n-1)}) \cdot \prob(L)} \cdot \odds(L) \\
 &= \Lambda_n \cdot \odds(L|D^{(n-1)}) \cdot \odds(\overline{L}) \cdot \odds(L) \\
 &= \Lambda_n \cdot \odds(L|D^{(n-1)})
\end{align*}

You may have noticed that the third step in the above deduction is only allowed if all $d_i$s are independent from each other. To be honest: they are not, but we have no idea how to account for it.

Moving on, we can turn this recursive formula into an iterative one:

\begin{equation}
\label{main_eq}
\odds(L | D^{(n)} ) =  \prod_{i=1}^n \Lambda_i  \cdot \odds(L)
\end{equation}

with $\odds(L)$ being the prior probability of an edge existing in the real world network.

\subsubsection{Avoiding computational restrictions}

Formula (\ref{main_eq}) is computationally inconvenient, because it can lead to numbers very close to zero, which might suffer from precision loss when stored in a binary floating-point format. It is more feasible to use their logarithms in this case. Instead of storing the two possible solutions to $\Lambda_i$, we can simply store the two solutions of $\ln(\Lambda_i)$ for each dataset. Then we can turn equation (\ref{main_eq}) into:

$$\ln(\odds(L|D^{(n)})) = \sum_{i=1}^n \ln (\Lambda_i) + \ln(\odds(L)) =: K$$

This function yields a log-odds, which we have called $K$. In order to turn the log-odds $K$ back into a probability, we can use the logistic function:

$$\prob(L | D^{(n)}) = \frac{e^K}{1 + e^K}$$

If we use the integration method described above, it yields global assessment of the probabilities of edge existence over all potential edges. This can be described as the map
$$ \p: \binom{V}{2} \rightarrow [0,1] $$

\subsection{Finding suitable error rate parameters}

The method described above still relies on the knowledge of values for $\alpha_i$ and $\beta_i$. However these numbers are not known to us, since we have no knowledge of the true network.

One way to overcome this problem would be to start with rough estimates on these parameter values, and use them to create a candidate integrated network. Then one could iteratively compute better parameter values and resulting integrated networks. This step could be repeated until a reasonable network is found. However, the produced series of networks cannot be expected to converge to the global maximum.

%\clearpage
%\subsubsection{Expectation maximization}
%To apply the EM algorithm approach, we must first find a way to express the expectation of our expression with respect to the true $G$:

%\begin{align*}
%  \expect_G \left [ \log \prob(X|\theta,G) \middle\vert X, \theta \right ] 
%  &= \expect_G \left [ 
%    \log \prod_{i=1}^n \prod_{e \in \binom{V}{2}} \prob(X_i^e|\theta_i,G) 
%  \middle\vert X, \theta \right ]\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \expect_G [ \log \prob(X_i^e|\theta_i,G) | X, \theta ]\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \expect_G [ \log (\Theta_{X,G,\theta}(i,e)) | X, \theta ]\\
%\end{align*}

%Let $\Gamma$ be the set of all possible graphs with $v$ nodes. Then, applying the definition of expectation to the above equation we can say:

%$$\expect_G [ \log (\Theta_{X,G,\theta}(i,e)) | X, \theta ] = \sum_{G \in \Gamma} \log(\Theta_{X,G,\theta}(i,e)) \cdot \prob(G|X,\theta) $$

%Now let $\Gamma^{(e+)} \subset \Gamma$ be the set of all graphs that contain $e$ and let $\Gamma^{(e-)} \subset \Gamma$ be the set of all graphs that \textbf{do not} contain $e$. Furthermore let
%$$ \Theta_{X,\theta}^{(+)}(i,e) = 
%\begin{cases}
%  e \in E_{X_i}: & 1- \beta_i\\
%  e \notin E_{X_i}: & \beta_i
%\end{cases}
%$$
%and 
%$$ \Theta_{X,\theta}^{(-)}(i,e) = 
%\begin{cases}
%  e \in E_{X_i}: & \alpha_i\\
%  e \notin E_{X_i}: & 1-\alpha_i
%\end{cases}
%$$
%Then

%\begin{align*}
%  \expect_G & \left [ \log \prob(X|\theta,G) \middle\vert X, \theta \right ] \\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \sum_{G \in \Gamma} \log(\Theta_{X,G,\theta}(i,e)) \cdot \prob(G|X,\theta)\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \left ( \sum_{G \in \Gamma^{(e+)}} \log(\Theta_{X,\theta}^{(+)}(i,e)) \cdot \prob(G|X,\theta) \right )
%     + \left ( \sum_{G \in \Gamma^{(e-)}} \log(\Theta_{X,\theta}^{(-)}(i,e)) \cdot \prob(G|X,\theta) \right )\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \left (\log(\Theta_{X,\theta}^{(+)}(i,e)) \sum_{G \in \Gamma^{(e+)}}  \prob(G|X,\theta) \right )
%     + \left (\log(\Theta_{X,\theta}^{(-)}(i,e)) \sum_{G \in \Gamma^{(e-)}}  \prob(G|X,\theta) \right )\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \log(\Theta_{X,\theta}^{(+)}(i,e)) \cdot \expect_G[\mathbb{I}(e \in E_G) | X,\theta] 
%     + \log(\Theta_{X,\theta}^{(-)}(i,e)) \cdot \expect_G[\mathbb{I}(e \notin E_G) | X,\theta]\\
%  &= \sum_{i=1}^n \sum_{e \in \binom{V}{2}}
%     \log(\Theta_{X,\theta}^{(+)}(i,e)) \cdot \p(e) 
%     + \log(\Theta_{X,\theta}^{(-)}(i,e)) \cdot (1-\p(e)) 
%\end{align*}

%Let's split up the sum according to those experimental graphs $X_i$ that contain an edge or do not contain an edge:

%\begin{align*}
%   \sum_{e \in \binom{V}{2}} & \left ( 
%     \sum_{X_i \in X^{(e+)}} 
%     \log(1-\beta_i) \cdot \p(e) 
%     + \log(\beta_i) \cdot (1-\p(e)) 
%     \right ) \\
%     &+ \left (
%     \sum_{X_i \in X^{(e-)}}
%     \log(\alpha_i) \cdot \p(e) 
%     + \log(1-\alpha_i) \cdot (1-\p(e))
%     \right )
%\end{align*}

%Next we want to find the maximum for $\theta = ((\alpha_1,\beta_1), \ldots, (\alpha_n,\beta_n))$. We'll have to do this for every $i$ separately, but fortunately, maximizing each member of the sum should also maximize the sum.

%Looking at the function itself it seems that if the average $p(e)$ over all edges is greater than 0.5, then $\beta_i = 0$ and $\alpha_i = 1$ maximizes the term. If the average $p(e)$ over all edges is less than 0.5, then $\beta_i = 1$ and $\alpha_i = 0$ maximizes the term. Now that is not helpful at all. Because we know the optimal value should be a bit more fine-tuned than that.

%But maybe I'm wrong, so let's try to find the partial derivatives for the $i$th $\alpha$ and $\beta$.

%\begin{align*}
%  \pardiff{\alpha_i} 
%  \sum_{e \in \binom{V}{2}} & \left ( 
%     \sum_{X_i \in X^{(e+)}} 
%     \log(1-\beta_i) \cdot \p(e) 
%     + \log(\beta_i) \cdot (1-\p(e)) 
%     \right ) \\
%     &+ \left (
%     \sum_{X_i \in X^{(e-)}}
%     \log(\alpha_i) \cdot \p(e) 
%     + \log(1-\alpha_i) \cdot (1-\p(e))
%     \right )
%\end{align*}

%\paragraph{Case: $e \in E_{X_i}$}

%\begin{align*}
%  \sum_{e \in \binom{V}{2}} & \left ( 
%     \sum_{X_i \in X^{(e+)}} 
%     \pardiff{\alpha_i} 
%     \log(1-\beta_i) \cdot \p(e) 
%     + \pardiff{\alpha_i} 
%     \log(\beta_i) \cdot (1-\p(e)) 
%     \right ) \\
%  &= 0
%\end{align*}

%\paragraph{Case: $e \notin E_{X_i}$}

%\begin{align*}
%  \sum_{e \in \binom{V}{2}} & \left ( 
%     \sum_{X_i \in X^{(e-)}} 
%     \pardiff{\alpha_i} 
%     \log(\alpha_i) \cdot \p(e) 
%     + \pardiff{\alpha_i} 
%     \log(1- \alpha_i) \cdot (1-\p(e)) 
%     \right ) \\
%  &= \sum_{e \in \binom{V}{2}} 
%     \pardiff{\alpha_i} 
%     \log(\alpha_i) \cdot \p(e) 
%     + \pardiff{\alpha_i} 
%     \log(1- \alpha_i) \cdot (1-\p(e)) \\
%  &= \sum_{e \in \binom{V}{2}} 
%     \frac{p(e)}{\alpha_i} + \frac{1-p(e)}{1-\alpha_i}
%\end{align*}

%Combining both cases:
%\begin{align*}
%  \sum_{e \in \binom{V}{2} \setminus E_{X_i}} & 
%    \frac{p(e)}{\alpha_i} + \frac{1-p(e)}{1-\alpha_i}\\
%  &= \frac{|\binom{V}{2} \setminus E_{X_i}| \cdot \alpha_i}{\alpha_i - \alpha_i^2} 
%    \sum_{e \in \binom{V}{2} \setminus E_{X_i}} p(e)
%\end{align*}

%Unfortunately, there is no $\alpha_i$ for which this term equals zero. (Unless all p(e) are zero, but that is not the point). For $\alpha_i = 0$ and $\alpha_i = 1$ the term is not even defined.

%Maybe I should look into MCMC instead.

To avoid this problem we use a Markov Chain Monte Carlo approach. The idea would be to alternately generate \textit{samples} for our error rate vector $\theta$ and for a potential true background graph $G$. This will result in a random walk. The better fit a particular state is the more likely it is to be assumed during the random walk. Thus, averaging over all intermediate steps should yield the best fitting parameter configuration. This approach poses five problems:
\begin{enumerate}
  \item Choosing an initial error rate vector $\theta_{\text{init}}$
  \item Sampling a potential true graph $G$, based on a given error rate vector $\theta$
  \item Sampling an error rate vector $\theta$ based on a true graph $G$
  \item Determining the likelihood of the current state $(\theta,G)$
  \item Averaging over random walk states.
\end{enumerate}

\subsubsection{Choosing an initial error rate vector $\theta_{\text{init}}$}

For now we will sample the initial values for $\alpha$ and $\beta$ from the uniform distribution $\mathcal{U}_{[0,1]}$. This also makes sense in the light of the details discussed in subsection \ref{sampling_theta}, since $\mathcal{U}_{[0,1]} = Be(1,1)$.

\subsubsection{Sampling a potential true graph $G$, based on a given error rate vector $\theta$}

Given an error rate vector $\theta$ we can use the integration method discussed above to infer posterior existence probabilities for each edge. We can use these probabilities to sample a potential $G$ by lookup. For each potential edge, we sample a $\mathcal{U}_{[0,1]}$ random number. If that random number is smaller than the posterior existence probability of that edge, we say exists. Otherwise we say it does not exist.

\subsubsection{Sampling an error rate vector $\theta$ based on a true graph $G$}
\label{sampling_theta}

To find a suitable distribution from which to sample these values, we should have a look at their nature. The error rates $\alpha$ and $\beta$ behave like the success rates for misreading each potential edge. So if we model each observation event over a potential edge as a Bernoulli experiment with such a success rate, the number of False Positives and False Negatives in an experimental graph $X_i$ would follow a binomial distribution. So how can we sample the success rates for a binomial distribution? Such is a accomplished using the Beta distribution. The Beta distribution takes two parameters, $a$ and $b$. $a$ and $b$ can be interpreted as one plus the number of successes and failures in a binomial distributed data set.

So if we compare $X_i$ to our currently assumed true graph $G$ and use it to count the number of True Positives (tp), False Positives (fp), True Negatives (tn) and False Negatives (fn), we can sample $alpha$ and $beta$ as follows:
\begin{align*}
  \alpha_i &\sim Be(\text{fp}_i+1, \text{tn}_i+1)\\
  \beta_i &\sim Be(\text{fn}_i+1, \text{tp}_i+1)
\end{align*}

\subsubsection{Determining the likelihood of the current state $(\theta,G)$:}

The likelihood of the current state can be computed using the posterior

$$\pi(\theta,G|X)$$
We cannot determine this directly. But we know that 
$$ \pi(\theta,G|X) = \frac{\pi(\theta,G,X)}{\prob(X)} $$
So the posterior distribution $\pi(\theta,G|X)$ is proportional to the joint distribution $\pi(\theta,G,X)$. We can determine the joint distribution as follows:
$$ \pi(\theta,G,X) = \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G) $$
Hence we can say that $$ \pi(\theta,G|X) \propto \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G) $$
So we need to determine the prior distribution $\pi(G)$, the prior distribution $\pi(\theta)$ and the likelihood $\prob(X|\theta,G)$.

\paragraph{Prior probability of $G$}
\begin{align*}
  \pi(G) &= \prod_{e \in \binom{V}{2}} \pi(G_e) \\
         &= (1-q)^{\left |\binom{V}{2}\setminus E_G \right |} \cdot q^{|E_G|}
\end{align*}
where $q$ is the prior probability of an edge to really exist.

\paragraph{Prior probability of $\theta$}

As discussed above, we assume a $Be(1,1)$ distribution for our $\alpha$ and $\beta$ values.

The probability density function for the beta distribution is:
$$ f(x;\alpha,\beta) = \frac{x^{\alpha-1}(1-x)^{\beta-1}}{\int_0^1 u^{\alpha-1} (1-u)^{\beta-1}\, du} $$

Thus, for $\alpha = \beta = 1$ :
$$ \pi(\theta) = f(\theta,1,1) = \frac{\theta^{0}(1-\theta)^{0}}{\int_0^1 u^{0} (1-u)^{0}\, du} = 1$$

So we might as well ignore this factor.

\paragraph{Posterior probability of $\theta$ and $G$}

\begin{align*}
  \prob(X|\theta, G) &= \prod_{i=1}^n \prob(X_i|\theta_i,G)\\
  &= \prod_{i=1}^n \prod_{e \in \binom{V}{2}} \prob( \mathbb{I}(e \in E_{X_i})|\theta_i,G)\\
  &= \prod_{i=1}^n \prod_{e \in \binom{V}{2}} \begin{cases}
    e \in E_G: & 
    \begin{cases}
      e \in E_{X_i}: & 1-\beta_i\\
      e \notin E_{X_i}: & \beta_i
    \end{cases}
    \\
    e \notin E_G: & 
    \begin{cases}
      e \in E_{X_i}: & \alpha_i\\
      e \notin E_{X_i}: & 1-\alpha_i
    \end{cases}
  \end{cases}
\end{align*}

However, since we don't know $G$, we will have to marginalize $G$ out. There are two possible courses of action: Using the Expectation-maximization (EM) algorithm or using a Markov chain Monte Carlo (MCMC) approach.

\paragraph{Adjustment for binary floating-point representation}

Again, since the likelihood value can be expected smaller than our machine floating-point precision, we will have to apply a logarithm function.

\begin{align*}
  \log\pi(\theta,G|X) &\propto \log (\pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G))\\
    &= \log\pi(G)+\log\pi(\theta)+\log\prob(X|\theta,G)\\
    &= \log((1-q)^{\left |\binom{V}{2}\setminus E_G \right |} \cdot q^{|E_G|})
       + \log(1)\\
       &\hspace{4ex} + \log\left ( \prod_{i=1}^n \prod_{e \in \binom{V}{2}} \begin{cases}
				e \in E_G: & 
				\begin{cases}
				  e \in E_{X_i}: & 1-\beta_i\\
				  e \notin E_{X_i}: & \beta_i
				\end{cases}
				\\
				e \notin E_G: & 
				\begin{cases}
				  e \in E_{X_i}: & \alpha_i\\
				  e \notin E_{X_i}: & 1-\alpha_i
				\end{cases}
			\end{cases} \right )\\
		&= \left |\binom{V}{2}\setminus E_G \right | \cdot \log(1-q) + |E_G|\cdot \log q\\
		&\hspace{4ex} + \sum_{i=1}^n \sum_{e \in \binom{V}{2}} \begin{cases}
				e \in E_G: & 
				\begin{cases}
				  e \in E_{X_i}: & \log(1-\beta_i)\\
				  e \notin E_{X_i}: & \log\beta_i
				\end{cases}
				\\
				e \notin E_G: & 
				\begin{cases}
				  e \in E_{X_i}: & \log\alpha_i\\
				  e \notin E_{X_i}: & \log(1-\alpha_i)
				\end{cases}
			\end{cases}
\end{align*}

\subsubsection{Averaging over random walk states}

If we do not wish to store all intermediate results we can keep an incremental mean value for each $\alpha_i$ and $\beta_i$ value. This is achieved using the recursive definition of the mean. Let $\{a\}_{N+1}$ be a series of $N+1$ numbers, then

\begin{align*}
  \expect(\{a\}_{N+1}) &= \frac{1}{N+1}  \sum_{k=1}^{N+1} a_k \\
    &= \frac{N}{N+1} \frac{1}{N} \left ( \sum_{k=1}^{N} a_k + a_{N+1}\right )\\
    &= \frac{N}{N+1} \left ( \frac{1}{N}\sum_{k=1}^{N} a_k + \frac{1}{N} a_{N+1}\right )\\
    &= \frac{N}{N+1} \left ( \expect(\{a\}_{N}) + \frac{1}{N} a_{N+1}\right )\\
\end{align*}

\subsection{Pseudocode}

\fbox{\parbox{\textwidth}{\scriptsize \input{pseudocode.tex}}}

%\subsubsection{Post-processing: Finding the highest peak in the state distributions}

%We try to find the optimal configuration of $\theta$, which consists of $n$ $(\alpha_i,\beta_i)$ pairs. That's $2n$ numbers between 0 and 1. So we can say the optimal $\theta$ is a point in a $2n$-dimensional hypercube. Having recorded all the intermediate states of $\theta$ and their associated likelihoods, we can raster-scan the hypercube in a number of steps. In the first step we divide each dimension in to $r$ intervals of equal sizes (thus dividing the hypercube into $r^{2n}$ sectors. Now we sort our recorded values into these sectors, adding their associated likelihoods to the of the sector. In each subsequent step we choose the sector with the highest likelihood sum from the previous iteration and repeat the procedure (subdivide and scan).

%Thus, given $s$ recorded states from the random walk, performing $d$ iterations will take $O((s + r^{2n}) \cdot d)$, reaching a precision of $r^{-d}$. As an example, given $n=5$ datasets, $s=10,000$ records, and interval size $r=2$, going for $d=100$ iterations takes 10,102,400 computing operations and reaches a precision of approximately $10^{-30}$ (which is smaller than binary floating point precision).

%\remark{\textbf{Problem:} Adding up the likelihoods poses a problem, since they can only be stored as logarithms. Adding up logarithms of course would be wrong ($\ln a + \ln b \neq \ln(a+b)$). So we need to find some way of performing this operation without converting them back into numbers smaller than our machine precision.}

%One way to overcome this problem would be to start with rough estimates on these parameter values, and use them to create a candidate integrated network. Then one could iteratively compute better parameter values and resulting integrated networks. This step could be repeated until a reasonable network is found. However, this idea suffers from two problems. (i) The produced series of networks cannot be expected to converge and (ii) a quantitative measure of the quality of would be required to assess whether a proposed solution is good or bad.
%In the next section we will address the second problem.

%\subsubsection{Assessing the likelihood of an integrated network model}

%Given our experimentally measured networks $X$, let $\theta = \theta_1, \ldots \theta_n$ be a series of vectors containing estimated error rate parameters, where 
%$$\theta_i = (\alpha_i, \beta_i) $$
%for each network $X_i$.

%If we use the integration method described above, it yields global assessment of the probabilities of edge existence over all potential edges. This can be described as the map
%$$ \p: \binom{V}{2} \rightarrow [0,1] $$
%Then we can write down a complete model of network integration as 
%$$ \model = (\p, \{\theta_i\}_{1 \leq i \leq n}) $$

%With these basic definitions we can now think about the likelihood of our model, given our experimental networks:

%\begin{align*}
%\likeli(\model | \{X_i\}_{1 \leq i \leq n}) &= \prob(\{X_i\}_{1 \leq i \leq n}| \model) = \prob \left (\bigwedge_{i=1}^n X_i | \model \right )\\
%  &= \prod_{i=1}^n \prob(X_i | \model)\\
%  &= \prod_{i=1}^n \prod_{e \in \binom{V}{2}} \prob(\mathbb{I}(e \in E_{X_i}) | \p(e) \wedge \theta_i)
%\end{align*}

%Again, here we step into the trap assuming that our experimental datasets are independent. But if we ignore that flaw for now we see that we end up with the product over all datasets for the product over all potential edges for the probability of observing the existence or non-existence, respectively, the selected edge, given our probability assessment for that edge and given our error rates for the selected dataset. Now what does that mean?

%\begin{center}
%	\begin{pspicture}(8,5)
%		\psline{->}(4,5)(2,3)
%		\rput(2.5,4){$\p(e)$}
%		\rput(2,2.8){$e \in E_G$}
%		
%		\psline{->}(4,5)(6,3)
%		\rput(5.8,4){$1-\p(e)$}
%		\rput(6,2.8){$e \notin E_G$}
%		
%		\psline{->}(2,2.6)(1,1)
%		\rput(1,2){$1-\beta_i$}
%		\rput(1,0.8){$e \in E_{X_i}$}
%		
%		\psline{->}(2,2.6)(3,1)
%		\rput(3,2){$\beta_i$}
%		\rput(3,0.8){$e \notin E_{X_i}$}
%		
%		\psline{->}(6,2.6)(5,1)
%		\rput(5,2){$\alpha_i$}
%		\rput(5,0.8){$e \in E_{X_i}$}
%		
%		\psline{->}(6,2.6)(7,1)
%		\rput(7,2){$1-\alpha_i$}
%		\rput(7,0.8){$e \notin E_{X_i}$}
%	\end{pspicture}
%	
%\end{center}

%If we follow the decision tree above we can derive the probability of experiment $X_i$ observing existence or non-existence of an edge based on $\p(e)$ and $\theta_i=(\alpha_i , \beta_i)$.

%$$
%\prob(\mathbb{I}(e \in E_{X_i}) | \p(e) \wedge \theta_i) =
%\begin{cases}
%	e \in E_{X_i} & \p(e)(1-\beta) + \alpha(1-\p(e)) \\
%	e \notin E_{X_i} & \beta \p(e)  + (1-\p(e))(1-\alpha)
%\end{cases}
%$$

%We summarize:
%$$
%\likeli(\model | \{X_i\}_{1 \leq i \leq n}) = \prod_{i=1}^n \prod_{e \in \binom{V}{2}}
%\begin{cases}
%	e \in E_{X_i} & \p(e)(1-\beta) + \alpha(1-\p(e)) \\
%	e \notin E_{X_i} & \beta \p(e)  + (1-\p(e))(1-\alpha)
%\end{cases}
%$$

%\subsubsection{Convergence to maximal likelihood}

%The second problem about the algorithm proposed earlier is to achieve convergence. We can currently think of two methods that could solve this problem:
%\begin{description}
%	\item[EM algorithm] An expectation maximization algorithm could be used. However, that method may only find local maxima.
%	\item[MCMC algorithm] such as Gibbs sampling. 
%\end{description}

%More here...

\end{document}

### 2011-08-19 src/main/latex/evaluation/evaluation.tex
Porting to mercurial

\documentclass{scrartcl}

\usepackage[british]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}  %theorem definition dependency
%\usepackage{moreverb} %source code listing dependency
%\usepackage{textcomp} %source code listing dependency
%\usepackage{listings} %source code listing dependency
\usepackage{color}    %source code listing dependency
\usepackage{titling}  %pdf metadata dependency
\usepackage{ifpdf}    %pdf metadata dependency
%\usepackage{graphicx}
%\usepackage{caption}
%\usepackage{pstricks}

%% DOCUMENT METADATA %%

\title{A probabilistic integration test harness}
\author{Jochen Weile, M.Sc.}
\date{\today}

%% LAYOUT OPTIONS %%
\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

%% CONFIGURE CODE HIGHLIGHTING %%
%\lstset{
%	tabsize=4,
%	language=Java,
%  basicstyle=\footnotesize,
%  upquote=true,
%  aboveskip={1.5\baselineskip},
%  columns=fixed,
%  showstringspaces=false,
%  extendedchars=true,
%  breaklines=true,
%  prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
%  frame=none,
%  showtabs=false,
%  showspaces=false,
%  showstringspaces=false,
%  identifierstyle=\ttfamily,
%  keywordstyle=\bf\color[rgb]{0.5,0,0.7},
%  commentstyle=\color[rgb]{0.5,0.5,0.5},
%  stringstyle=\color[rgb]{0,0,1},
%}

%% SUPPORT FOR PDF METADATA %%
\ifpdf
	\pdfinfo {
		/Author (\theauthor)
		/Title (\thetitle)
		/CreationDate (D:20110216120000)
	}
\fi

%% DEFINE MACROS %%
%\newcommand{\prob}{\mathbb{P}}
%\newcommand{\odds}{\mathbb{O}}
%\newcommand{\likeli}{\mathcal{L}}
%\newcommand{\model}{\mathcal{H}}
%\newcommand{\p}{\text{p}}
%\newcommand{\expect}{\mathbb{E}}
%\newcommand{\pardiff}[1]{\frac{\partial}{\partial #1}}
\newcommand{\remark}[1]{\begin{center}\fcolorbox{black}{yellow}{\parbox{.8\textwidth}{#1}}\end{center}}

%% DEFINE ENVIRONMENTS %%
%\newtheorem{theorem}{Theorem}
%\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\section{Requirements}
\subsection{Workflow model}

We aim at providing a software solution for testing the qualities of network integration algorithms. To this end we propose a simple workflow model. Each workflow cycle begins with the generation of a random graph which is assumed as the true background network. From this true model we derive experimental measurements, which serve as input for the integration algorithm to be tested. The integrated network computed by the integration method is then compared to the original true network.

In order to establish the average behaviour of the integration method it would be desirable to perform not only one but many workflow cycles and then determine the mean and variance over the results. Furthermore we would like to examine algorithms given varying parameters, so the workflow model should allow for the iteration over variable parameters. However, performing a large number of workflow cycles while varying over multiple parameters can result in significantly long runtime when performed in serial, especially if the integration method itself is computationally expensive. Thus, support for parallelization is necessary.

\subsection{Extensibility}

For almost all steps of a workflow cycle, many different applications are possible. There are different classes of graphs which are possible as true background graphs, such as random graphs or scale-free graphs as well as different algorithms for generating them. Similarly, for the experiment simulation and result evaluation steps there are different solutions imaginable, all of which have different advantages and dis-advantages. Finally, there are many different integration methods one may want to test. To keep the test harness open to new ideas regarding these components it would be useful to employ a plug-in architecture. New components can then be implemented as plug-ins and deployed without the need of adapting the core system.

\subsection{Random number generation}

A substantial part of the test harness will depend on random number generation, such as the graph populators, the experiment simulation and some of the integration methods (most importantly the MCMC method). In order to minimize potential side effects like correlation between the the numbers generated in the different components, we will require a centralized random engine which can be accessed queried from all components. Since multiple workflow cycles may be performed in parallel, this random engine needs to be thread-safe, while at the same time not introducing a bottleneck to the parallelization.

\section{Implementation}

\subsection{Input}

Since the input for the test harness consists of a complicated test protocol containing descriptions of the different steps of the workflow plus the definitions of potential parameter variables, we decided to use custom XML documents to convey this input. We specified an XML schema definition which governs the syntax of these documents. A description of the input format can be found on the documentation page\footnote{http://bio-nexus.ncl.ac.uk/projects/harness}. Here, we will give a short overview of the contents of a typical input XML file.%TODO: write xml description on site

The first important section of the input protocol are variable definitions. Each variable has an id, a type and a range definition. The ids are used to establish a way of referencing a variable later on in the workflow. At the time this document is written, there are two types of variables: (i) incremental variables, which require a start value, end value and skip size in their range definition, and (ii) exponential variables, which require a start value, an exponent and a number of iterations as their range definition.

The next section of the protocol defines which graph implementation to use. The Harness core package only comes with one graph implementation: the map graph. Details about the map graph are given in section \ref{graph model}. However, Harness offers an open graph API, so users can implement and use their own graph models. The XML format also allows custom parameters to be passed to graph implementations. These properties can either be hard-coded values or references to the variables defined in the previous section.

Another section in the protocol is used to define which graph population model to use when simulating true biological graphs. The Harness only contains one population model: The scale-free population model (See section \ref{graph population}). Other population models can be implemented and applied using the Harness API. Custom parameters can also be passed to graph models. For example, the scale-free graph model requires a number of seed nodes and a size parameter.

The final section in the protocol is the description of the integration process. It determines which integration method will be tested by the harness and how many times the workflow should be executed. New integration methods in addition to the ones already contained in the Harness core package can be implemented and deployed using the API. The integration description also contains two sub-sections, describing the simulation of Gold Standards and experiments. Both are used in the same way. First, the experiment type is given. As with other plug-ins experiment types can also be implemented against the Harness API. Then, the number of experiments is given. And finally, custom properties can be passed to the experiment simulation. For example, the `simple experiment' simulation requires sensitivity and specificity parameters. These parameters, as well as the number of replicas can be hard-coded values or references to variables.

\subsection{Workflow}

The workflow controller iterates over all possible combinations of variables defined in the protocol. In each iteration the controller will execute the pre-determined number of workflow replicas in a concurrent fashion. That is, multiple workflow instances will be run simultaneously, the number of which is defined by the user and is optimally chosen to match the number of available CPU cores.

Each of the workflow threads will begin instanciating and populating a new graph, which will serve as the simulated `true' graph. This graph will then be passed to the appropriate number of experiment simulations. The resulting simulated evidental graphs will then be given to the integration method in question. Finally, the integration method's output will be compared to the original `true' graph and the results are reported back to the workflow controller, which writes them to the output file.

\subsection{Graph model}
\label{graph model}
The standard graph model used in the Harness is the map graph, which uses hash maps to store the graph's topology. This way of representing graphs is usually the most memory-efficient, when applied to sparse graphs with less then 8\% of possible edges. Molecular biological networks fall into this category due to their scale-free topology.

\subsection{Graph population}
\label{graph population}

The standard graph population model used in the Harness is the preferential attachment model for scale-free graphs. The preferential attachment algorithm begins with number of fully connected seed nodes. In each following iteration, the algorithm adds a new node to the graph and connects this node to another random pre-existig node. Here, the probability of a node to be chosen for a new connection is directly proportional to its degree.

%TODO: Pseudocode?

\section{Experiment simulation}

%TODO: fast experiment + pseudocode

\section{Integration algorithms}

\subsection{Evaluation methods}

\end{document}

### 2011-08-19 src/main/latex/paper/paper.tex
Porting to mercurial

\documentclass{bioinfo}
\copyrightyear{2011}
\pubyear{2011}

\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{color}

%% DEFINE MACROS %%
%% Math macros %%
\newcommand{\prob}{\mathbb{P}}
\newcommand{\odds}{\mathbb{O}}
\newcommand{\likeli}{\mathcal{L}}
\newcommand{\p}{\text{p}}
\newcommand{\expect}{\mathbb{E}}
%% Bio macros %%
\newcommand{\species}[1]{\textit{#1}}
%% Layout macros %%
\newcommand{\remark}[1]{\begin{center}\fcolorbox{black}{yellow}{\parbox{.8\textwidth}{#1}}\end{center}}
\newcommand{\note}[1]{{\color{red}[#1]}}
\newcommand{\REF}{{\color{red}[REF]}}

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% CONTENTS %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\firstpage{1}

\title[Bayesian integration of networks without gold standards]{Bayesian integration of networks without gold standards}

\author[J. Weile \textit{et~al.}]{Jochen Weile\,$^{1}$, Katherine James\,$^{1}$, Jennifer Hallinan\,$^{1}$, Simon J Cockell\,$^{2}$, Phillip Lord\,$^{1}$, Anil Wipat\,$^{1,3}$ and Darren Wilkinson\,$^{3,4,}$\footnote{to whom correspondence should be addressed}}

\address{$^{1}$School of Computing Science, $^{2}$Bioinformatics Support Unit, $^{3}$Centre for Integrative Systems Biology of Ageing and Nutrition, $^{4}$School of Mathematics and Statistics, Newcastle University, Newcastle upon Tyne NE1 7RU, United Kingdom}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}
\section{Motivation:}
Biological experiments allow insights into networks of processes inside a cell, but are subject to errors. However, thanks to the overlap between the large number of experiments reported in public databases it is possible to assess the chances of individual observations to be correct. In order to do so, existing methods rely on high-quality ``gold standard'' reference networks, but such reference networks are not always available.

\section{Results:}
We present a novel algorithm for computing the probability of experimental observations to be correct that operates without gold standard reference data. We show that our algorithm even outperforms existing gold standard based methods. Finally, we apply the new algorithm on a large collection of genetic interaction and protein-protein interaction experiments.

\section{Availability:}
The integrated dataset and a reference implementation of the algorithm as a plug-in for the Ondex data integration framework are available for download at \href{http://bsu.ncl.ac.uk/nogold}{http://bsu.ncl.ac.uk/nogold}.

\section{Contact:} \href{darren.wilkinson@ncl.ac.uk}{darren.wilkinson@ncl.ac.uk}
\end{abstract}

\section{Introduction}

A significant portion of knowledge about molecular biological processes is distributed over a large number of online databases \citep{stein_creating_2002}. This knowledge has been obtained through experiments performed in laboratories all over the world. Overlaps often exists across the contents of these databases. The sub-discipline of integrative bioinformatics aims at collating this knowledge and making it accessible to both humans and computers. A large number of approaches to this problem exists. %\note{Mention some}.

A popular integration paradigm is the construction of functional networks \citep{von_mering_string:_2003,warde-farley_genemania_2010}\note{Also cite Katherine}. Functional networks represent several different types of relationships between biological entities in an abstract manner. Associations such as genetic interactions, protein-protein interactions, gene regulation and co-expression, are combined into abstract statements of functional relatedness, which are termed functional interactions.

An alternative paradigm is semantic data integration \citep{cheung_yeasthub:_2005,smith_obo_2007,koehler_graph-based_2006}. These approaches aim at representing the biological information (and as much of its meaning as possible) in a computationally accessible fashion. Rather than generalizing over all types of associations between entities to infer functional interactions, each type of association is considered separately.

An important question in data integration is how to assess the degree of confidence in each statement, that is, how likely the statement is to be correct. Several popular solutions to this problem exist in the context of functional networks \citep{lee_probabilistic_2004,troyanskaya_bayesian_2003}.
These methods assess the quality of each input dataset against one or more additional datasets of higher quality, usually manually-curated collections. Based on the confidence measures gained from this comparison it is then possible to calculate a confidence measure for each functional interaction. The high-quality datasets used in these comparisons are often referred to as `gold standards'.

Inferring confidence assessments for semantically integrated datasets is challenging, because each single type of association must be scored separately. However, gold standards only exist for some types of associations. Therefore, it is desirable to find a method that can infer confidence measures on biological networks in the absence of a gold standard, that is, based only on the existing experimental data. So far, only partial solutions to this problem exist. These solutions infer confidence measures for a specific type of association only, for example, protein-protein interactions \citep{bader_gaining_2004}.\note{Find more}

%Whether or not this description is justified is a matter of dispute \REF.

%It would be desirable to investigate these ideas at a more detailed level, that is, moving on from the question: ``How likely are two given proteins to be functionally related?'', to more specific questions such as: ``How likely is it that two given proteins physically interact?''. However, at this level it becomes infeasible to rely on gold standards, \note{since the level of completeness and quality required has not yet been achieved.} simply because none exist \REF. Therefore, it is necessary to develop a method that can infer confidence measures for semantic statements about molecular biological entities, based only on the existing data.

To provide a more generic solution to this problem, we present a fully Bayesian method that calculates, for each statement in a semantically-integrated dataset, the probability that it is true. We have evaluated the method's effectiveness in comparison to related methods. The validity of any results of the method's application to real data is hard to verify without knowing the absolute biological truth. Therefore, we have developed a tool that tests integration methods on simulated data with the same characteristics as real biological networks.

\begin{methods}
\section{Methods}

\subsection{Probabilistic integration}

The complete system of processes of a certain type within the cell (e.g. protein-protein interactions) can be modelled as a network $G=(V,E)$, where entities, such as proteins, are nodes (vertices) $V=\{v_1,\ldots,v_N\}$ and their associations are edges $E = \{e_1, \ldots,e_M\} \subseteq \binom{V}{2}$. If one considers each pair of entities to potentially have an edge, it is possible to model the process of prediction of such an edge by an experiment as follows. Let $X = \{X_1, \ldots, X_n \}$ be a vector of $n$ networks, which have been experimentally derived from $G$. Considering a single potential edge $e$, each experiment $X_i$ makes a statement about $e$'s existence. Let $D^e_i$ be a random variable that assumes realisation 1 when the $i$th experiment $X_i$ predicts that the edge exists and 0 when it predicts that the edge does not exist. Let $d^e_i$ be the measured realisation from $X_i$, then $(D^e_i=d^e_i)$ is the event that the measured realisation in experiment $i$ is taking place. Furthermore, let $D^e_{(n)}$ be the vector of all $n$ experimental measurement events $(D^e_i=d^e_i)$ for the edge. Finally, let $L^e$ be the event that the edge really does exist in $G$. We are interested in $\prob(L^e|D^e_{(n)})$, that is, the probability that the edge really exists given all our experimental measurements.

An important concept necessary to determine this probability is the Bayes factor \citep{goodman_toward_1999}. For each of the $n$ experiments a Bayes factor $\Lambda_i$ can be determined, which is defined as follows:

\begin{equation} 
	 \Lambda_i := \frac{\prob(D^e_i=d^e_i|L^e)}{\prob(D^e_i=d^e_i|\neg L^e)} 
\end{equation}

If experiment $i$ predicts that the edge exists, then $\Lambda_i$ is the probability of a true positive in $i$ divided by the probability of a false positive in $i$. Otherwise, if $i$ predicts that the edge does not exist, then $\Lambda_i$ is the probability of a false negative in $i$ divided by the probability of a true negative in $i$.

Then, under the assumption that all measurements are independent from each other, the Bayes theorem can be used to show:

\begin{equation} 
	\odds(L^e|D^e_{(n)}) = \Lambda_n \cdot \odds(L^e|D^e_{(n-1)})
\end{equation}

A full proof of the above equation is provided in the supplementary material. This recursive equation can be expressed iteratively as follows:

\begin{equation} 
	\odds(L^e|D^e_{(n)}) = \prod_{i=1}^n \Lambda_i \cdot \odds(L^e)
\end{equation}

That is, the odds of the edge to exist given all the experimental measurements is the product of the Bayes factors for these measurements with the prior odds of edge existence. These odds can be converted into the equivalent probability.

As mentioned above, the calculation assumes that experimental measurements are stochastically independent. This assumption is not valid for real data, and thus introduces a potential source of error into the methodology. Lee and colleagues address this problem by introducing dependency co-efficients for their datasets, a solution which relies heavily on guesswork \citep{lee_probabilistic_2004}. Instead we focus on making our method robust towards this kind of error.

In order to calculate the Bayes factors above it is necessary to determine the rates of false positives and false negatives in each dataset. One approach is to compare each dataset to a gold standard and count the number of differences. However, due to previously discussed lack of gold standards, this approach is not feasible here. Therefore, the only available option is to evaluate the datasets against one another in the form of a common consensus.

A naive approach is to start with random values for the error rates, and to use them to create a candidate integrated network. Updated parameter values and the resulting integrated networks can then be iteratively computed. However, a series of networks produced by this method does not typically converge to any sensible result (data not shown).

To overcome this problem, a fully Bayesian approach was employed to generate samples from the full joint posterior distribution of $\pi(\theta,G|X)$, where $\theta = \{ (\alpha_1,\beta_1), \ldots , (\alpha_n,\beta_n) \}$ is the vector of error rates associated with the members of the vector of experimental networks $X$, and where $\alpha_i$ is the false positive rate of $X_i$ and $\beta_i$ is the false negative rate of $X_i$..
%This algorithm alternately generates samples of the error rates and of a potential true background graph. The method performs a random walk from one sampled state to another. The more accurate a solution that a particular state represents, the more likely it is to be assumed during the random walk. Thus, averaging over the error rate samples of all steps in the algorithm yields a usable parameter configuration.
%\note{Formalise as way of sampling from full joint posterior. Explain Bayesian inference and Gibbs approach}

%The vector of error rates associated with the members of the vector of experimental networks $X$, is denoted as $\theta = \{ (\alpha_1,\beta_1), \ldots , (\alpha_n,\beta_n) \}$, where $\alpha_i$ is the false positive rate of $X_i$ and $\beta_i$ is the false negative rate of $X_i$.

However, instead of sampling from $\pi(\theta,G|X)$ directly, we employ Gibbs sampling \citep{hastings_monte_1970} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.

The algorithm proceeds in cycles. At the beginning of each new cycle, a potential true graph $G$ needs to be sampled based on the error rate vector $\theta$. The sampling is accomplished by using the Bayesian method discussed above to infer posterior existence probabilities for each edge. These probabilities can then be used to sample a potential $G$ by lookup. That is, for each potential edge, a $\mathcal{U}_{[0,1]}$ random number is sampled. If that random number is smaller than the posterior existence probability of that edge, $G$ will contain the edge. Otherwise $G$ will not contain the edge.

The second step in each cycle is the sampling of a new error rate vector $\theta$ based on $G$. The error rates $\alpha_i$ and $\beta_i$ behave like the success rates for misreading each potential edge. Modelling each observation event over a potential edge as a Bernoulli experiment with such a success rate, the number of false positives and false negatives in an experimental graph $X_i$ would follow a binomial distribution. The success rates for a binomial distribution can be sampled using the Beta distribution. The Beta distribution takes two parameters, $a$ and $b$. The parameters $a$ and $b$ can be interpreted as one plus the number of successes and failures, respectively, in a binomial distributed dataset. In order to obtain values for $a$ and $b$, one can compare each $X_i$ to the currently assumed true graph $G$ and use it to count the number of supposed True Positives (tp), False Positives (fp), True Negatives (tn) and False Negatives (fn). Then, $\alpha$ and $\beta$ can be sampled as follows:
\begin{align}
  \alpha_i &\sim Be(\text{fp}_i+1, \text{tn}_i+1)\\
  \beta_i &\sim Be(\text{fn}_i+1, \text{tp}_i+1)
\end{align}

To initiate the algorithm as a whole, we need to generate initial error rate vector $\theta_{\text{init}}$. It is sufficient to sample the initial values for each $\alpha_i$ and $\beta_i$ from $Be(1,1)$, which is equivalent to the uniform distribution $\mathcal{U}_{[0,1]}$. Pseudocode for the complete algorithm is given in the supplementary material.

%Finally, a score analogous to the likelihood of the current state can be computed as $\pi(\theta,G|X)$.
%Lacking a way to determine this value directly, one may exploit the following equation.

%\begin{equation} 
%	 \pi(\theta,G|X) = \frac{\pi(\theta,G,X)}{\prob(X)} 
%\end{equation}

%Thus, the posterior distribution $\pi(\theta,G|X)$ is proportional to the joint distribution $\pi(\theta,G,X)$. The joint distribution may in turn be determined as follows.

%\begin{equation} 
%	 \pi(\theta,G,X) = \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G)
%\end{equation}

%In summary, the following statement can be made.

%\begin{equation} 
%	 \pi(\theta,G|X) \propto \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G) 
%\end{equation}

%As a consequence, three values need to be determined: the prior distribution $\pi(G)$, the prior distribution $\pi(\theta)$ and the likelihood $\prob(X|\theta,G)$.

\subsection{Evaluation method}

In order to evaluate the method described above and to compare it against other probabilistic integration methods, a simulation and testing environment was created. The testing tool creates a random graph according to a specified topology. In the simulation, this graph assumes the role of a true biological graph. The tool then derives a set of graphs from the true graph with pre-determined error rates. In the simulation these graphs assume the role of experimental datasets.

The simulated experimental datasets are subsequently passed to the integration method under investigation. The integrated graph resulting from the integration method is then compared with the original simulated true graph to evaluate the integration method's performance. Such a testing workflow can be programmed to be executed a large number of times in order to measure a method's average behaviour. Furthermore, the testing tool allows for the automatic variation of different input parameters.

The simulated true graph is created as a scale-free graph using a preferential attachment algorithm, since most molecular-biological graphs have been shown to be belong to the scale-free class \citep{jeong_lethality_2001, eisenberg_preferential_2003}. A description of the algorithm can be found in the supplementary material.
\note{Mention how this does not match the algorithm's definition of the prior distribution of $G$, thus making it more challenging for the algorithm to perform.}%TODO

The next step consists of the simulation of experimental measurements on the true graph. This is the most crucial step of the artificial testing environment as it is responsible for replicating all the different faults and problems of real data. The simplest type of error occurring in experimental measurements is random noise. This type of error is easily simulated by randomly inserting edges that do not exist in the real data and removing edges that do exist in the real data until the desired error rates are reached. The next problem is systematic error or experimental bias. This phenomenon in particular leads to the violation of stochastic independence between datasets. 
\note{TODO: Bias simulation has yet to be implemented.}%TODO <-- 
Finally, another type of error that easily escapes one's attention is the lack of negative records. Most experiments do not report which pairs of nodes do not possess an edge. As a consequence, it is hard to differentiate between node pairs that have been observed not to interact and node pairs that have not been subject to testing. 
\note{TODO: Low-throughput experiments have yet to be implemented}%TODO <--

After the simulated evidential graphs have been given to the integration method in question the resulting integrated network is evaluated against the original graph. There are a number of different quality measures that can be applied. One important aspect to measure the accuracy of the method to estimate the error rates in the individual experiments. We can define the quadratic loss regarding the error estimates as follows.
\begin{equation}
  L_\text{ER} = \frac{1}{2n}\sum_{i=1}^n (\alpha_i - \hat{\alpha_i})^2 + (\beta_i - \hat{\beta_i})^2
\end{equation}
where $\hat{\alpha_i}$ and $\hat{\beta_i}$ are the estimates of the false positive and false negative rates for experiment $i$ according to the integration method in question.

Also, to measure the accuracy of the final edge probabilities produced by the method, we can define further loss functions. Since we can expect a vast number of true negatives when working with sparse, scale-free graphs, it would be helpful to see the loss over interacting and non-interacting node pairs separately. These can be interpreted as analogue to the algorithms false negative rate and false positive rate.

\begin{align} 
	L^{(+)} &= \frac{1}{|E_G|}\sum_{e \in E_G} (1 - \prob(e \in E_G|X))^2\\
	L^{(-)} &= \frac{1}{|\binom{V}{2} \setminus E_G|}\sum_{e \in \binom{V}{2} \setminus E_G} \prob(e \in E_G|X)^2
\end{align}
Other performance metrics such as the area under ROC curves or weighted sums over the quadratic loss functions above are also possible to apply. However we find that the separate loss functions shown above give a more detailed picture of a method's behaviour.
\note{Maybe move to discussion. Discuss other possible scoring functions, e.g. weighted sums over FPs and FNs}

We have evaluated the new algorithm in comparison with two gold standard-based methods presented by Lee and colleagues \citep{lee_probabilistic_2004} and James and colleagues \REF. For our evaluation we have examined the following scenario: Each integration method is given the task of processing a number (3,5,7,9,11) of high-throughput (HTP) experiments with average false negative rate $0.15$ and average false discovery rate of $0.15$ (corresponds to false positive rate of 0.0006). In the case of the gold standard-based methods, one of the input experiments serves as the gold standard.

\subsection{Semantic integration}

In addition to the evaluation on artificial data it is important to analyse the method's performance on real biological data. To this end we have used the semantic data integration system Ondex \citep{koehler_graph-based_2006} to gather as much data as possible on the \species{Saccharomyces cerevisiae} protein-protein interaction (PPI) network. A new PSI-MI 2.5 XML \citep{kerrien_broadening_2007} parser was written to import data from the BioGRID \citep{breitkreutz_biogrid_2008}, MINT \citep{chatr-aryamontri_mint:_2007}, IntAct \citep{hermjakob_intact:_2004} and MIPS-MPACT \citep{guldener_mpact:_2006}. The resulting Ondex dataset represented proteins, their interactions, the experiments in which the interactions have been observed and the publications in which the experiments were described. A semantic merger method was developed to clean the dataset of its redundancies. Further details on the semantic integration procedure can be found in the supplementary material.

\end{methods}

\section{Results}

Figure \ref{htp} shows the methods' performances over different numbers of input HTP experiments. It is clearly visible that the new fully Bayesian method outperforms both gold standard-based methods, particularly given a low number of input datasets.

\begin{figure}[!tpb]
\centerline{\includegraphics[width=.5\textwidth]{fig1.pdf}}
\caption{Evaluation of the algorithm in comparison with GS-based methods by Lee and colleagues and James and colleagues given different numbers of HTP experiments. Top: Average loss over interacting node pairs ($L^{(+)}$). Centre: Average loss over non-interacting node pairs ($L^{(-)}$). Bottom: Average loss over regarding error rate estimates ($L_\text{ER}$).}
\label{htp}
\end{figure}

When given a mixed set of HTP and LTP experiments, ...\note{TODO: Implement LTP simulation}

\section{Discussion}

\note{TODO: Write discussion}

\section*{Acknowledgements}

\paragraph{Funding\textcolon} The authors are pleased to acknowledge funding from the Biotechnology and Biological Sciences Research Council (BBSRC) Systems Approaches to Biological Research (SABR) initiative (Grant number BB/F006039/1).

\paragraph{Contributions\textcolon} ...

\paragraph{Conflict of interest\textcolon} None declared.

\bibliographystyle{natbib}

\bibliography{paper}

\end{document}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/LoggingInit.java

@@ -1,0 +1,71 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness;
+
+import java.io.BufferedInputStream;
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.IOException;
+import java.io.InputStream;
+import java.util.logging.LogManager;
+import uk.ac.ncl.cs.harness.io.ResourcePaths;
+
+/**
+ * Logging system initialisation. Calling <code>LoggingInit.init();</code>
+ * will read the local logging.config file.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class LoggingInit {
+
+    /**
+     * Points to the logging config file.
+     */
+    public static final String FILE_NAME = "logging.config";
+
+    /**
+     * Reads the logging config file and applies its settings.
+     */
+    public static void init() {
+
+        try {
+
+            InputStream in = ClassLoader.getSystemResourceAsStream(FILE_NAME);
+            
+            if (in == null) {
+
+                System.err.println("WARNING: Unable to obtain logging "
+                        + "configuration file from classpath! Trying to "
+                        + "use IDE file...");
+
+                File file = new File(ResourcePaths.RESOURCES, FILE_NAME);
+                in = new FileInputStream(file);
+
+            }
+
+            LogManager logManager = LogManager.getLogManager();
+            logManager.readConfiguration(new BufferedInputStream(in));
+
+        } catch (IOException ex) {
+            throw new RuntimeException("Could not find logging configuration "
+                    + "file! Report this as a bug.",ex);
+        }
+
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/Main.java

@@ -1,0 +1,160 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness;
+
+import java.io.File;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import org.apache.commons.cli.CommandLine;
+import org.apache.commons.cli.CommandLineParser;
+import org.apache.commons.cli.HelpFormatter;
+import org.apache.commons.cli.Options;
+import org.apache.commons.cli.ParseException;
+import org.apache.commons.cli.PosixParser;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionRegistry;
+import uk.ac.ncl.cs.harness.io.OutputController;
+import uk.ac.ncl.cs.harness.io.ProtocolParser;
+import uk.ac.ncl.cs.harness.workflow.Protocol;
+import uk.ac.ncl.cs.harness.workflow.ConcurrentWorkflow;
+
+/**
+ * The integration test harness main class.
+ *
+ * @author Jochen Weile
+ */
+public class Main {
+
+    /**
+     * Main method.
+     * @param args expects the name of the protocol.xml file to be used.
+     */
+    public static void main(String[] args) {
+
+        LoggingInit.init();
+
+        try {
+
+            Main main = new Main();
+
+            main.run(main.processArgs(args));
+
+        } catch (Throwable t) {
+            logProcessedError(t);
+            Logger.getLogger(Main.class.getName()).log(Level.FINE, t.getMessage(), t);
+            System.exit(1);
+
+        }
+        
+    }
+
+    private static void logProcessedError(Throwable t) {
+        StringBuilder b = new StringBuilder(256);
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/exceptions/ConfigurationException.java

@@ -1,0 +1,61 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.exceptions;
+
+/**
+ *
+ * Indicates that the program has been configured incorrectly.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class ConfigurationException extends Exception {
+
+    /**
+     * Constructor with cause.
+     * @param cause The cause of the exception.
+     */
+    public ConfigurationException(Throwable cause) {
+        super(cause);
+    }
+
+    /**
+     * Constructor with message and cause.
+     * @param message A message.
+     * @param cause The cause of the exception.
+     */
+    public ConfigurationException(String message, Throwable cause) {
+        super(message, cause);
+    }
+
+    /**
+     * Constructor with message.
+     * @param message A message.
+     */
+    public ConfigurationException(String message) {
+        super(message);
+    }
+
+    /**
+     * Empty constructor.
+     */
+    public ConfigurationException() {
+    }
+
+    
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/exceptions/GraphException.java

@@ -1,0 +1,59 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.exceptions;
+
+/**
+ * Indicates that a problem has occurred within the graph.
+ * @author Jochen Weile, M.Sc.
+ */
+public class GraphException extends RuntimeException {
+
+    /**
+     * Constructor with cause.
+     * @param cause The cause of the exception.
+     */
+    public GraphException(Throwable cause) {
+        super(cause);
+    }
+
+    /**
+     * Constructor with message and cause.
+     * @param message A message.
+     * @param cause The cause of the exception.
+     */
+    public GraphException(String message, Throwable cause) {
+        super(message, cause);
+    }
+
+    /**
+     * Constructor with message.
+     * @param message A message.
+     */
+    public GraphException(String message) {
+        super(message);
+    }
+
+    /**
+     * Empty constructor.
+     */
+    public GraphException() {
+    }
+
+    
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/exceptions/GraphMixupException.java

@@ -1,0 +1,74 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.exceptions;
+
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+
+/**
+ * Indicates that an element belonging to one graph has been inserted
+ * into another, alien graph.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class GraphMixupException extends GraphException {
+
+    /**
+     * Empty constructor.
+     */
+    public GraphMixupException() {
+    }
+
+    /**
+     * Constructor with message.
+     * @param message A message.
+     */
+    public GraphMixupException(String message) {
+        super(message);
+    }
+
+    /**
+     * Constructor with message and cause.
+     * @param message A message.
+     * @param cause The cause of the exception.
+     */
+    public GraphMixupException(String message, Throwable cause) {
+        super(message, cause);
+    }
+
+    /**
+     * Constructor with cause.
+     * @param cause The cause of the exception.
+     */
+    public GraphMixupException(Throwable cause) {
+        super(cause);
+    }
+
+
+    /**
+     * Constructor with the alien node and the graph
+     * @param node The node which has been falsely inserted.
+     * @param graph The graph which the node was applied to.
+     */
+    public GraphMixupException(Node node, Graph graph) {
+        super("Node \""+node.getId()+"\" belongs to graph \""+
+                node.getOwningGraph().getName()+"\" but was used with graph \""+
+                graph.getName()+"\".");
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/experiments/Experiment.java

@@ -1,0 +1,53 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import uk.ac.ncl.cs.harness.extsupport.Configurable;
+import uk.ac.ncl.cs.harness.graph.Graph;
+
+/**
+ * <p>The experiment interface can be implemented as a possible extension to the
+ * harness. All implementing classes should carry the <code>@Extension</code> annotation
+ * as well as mark all their vital bean setters with <code>@ExtensionProperty</code>
+ * annotations.</p>
+ * 
+ * <p>An experiment models the process of a true biological graph being measured by
+ * a wet-lab experiment. A new graph is generated in the process, which contains
+ * certain errors as compared to the original graph. The details of how these errors
+ * are introduced depends on the implementation of the experiment.</p>
+ *
+ * <p>In order to apply an experiment to a given template graph, obtain an
+ * <code>Experiment</code> instance from the <code>ExtensionRegistry</code> and
+ * invoke the <code>perform(..)</code> method using the true graph and a new empty graph
+ * as parameters. the empty graph will then be filled according to the experiment's
+ * specifications.</p>
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public interface Experiment extends Configurable {
+
+    /**
+     * Performs the experiment on the true biolocial graph <code>trueGraph</code>
+     * The results are written into the given graph <code>emptyGraph</code>
+     * @param trueGraph the true biological template graph
+     * @param emptyGraph an empty graph which will be filled with the experiment's
+     * results.
+     */
+    void perform(Graph trueGraph, Graph emptyGraph);
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/experiments/FastExperiment.java

@@ -1,0 +1,200 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import java.util.ArrayList;
+import java.util.List;
+import java.util.logging.Logger;
+import cern.jet.random.Poisson;
+import cern.jet.random.Uniform;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ * A basic implementation of the <code>Experiment</code> interface. The experiment
+ * introduces errors according to given sensitivity and FP-rate values.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Extension(id="fastexperiment")
+public final class FastExperiment implements Experiment {
+
+    /**
+     * The experiment's sensitivity.
+     */
+    private double sensitivity;
+
+    /**
+     * The experiment's false positive rate.
+     */
+    private double fpr;
+
+    /**
+     * A poisson distributed random number generator.
+     */
+    private Poisson poisson;
+    /**
+     * A uniform random number generator.
+     */
+    private Uniform uniform;
+
+    public FastExperiment() {
+//        MersenneTwister r = new MersenneTwister(RandomSeeds.getInstance().nextSeed());
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/experiments/ReductionistExperiment.java

@@ -1,0 +1,173 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import cern.jet.random.Poisson;
+import cern.jet.random.Uniform;
+import java.util.ArrayList;
+import java.util.List;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Neighbour;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="reductionist")
+public class ReductionistExperiment implements Experiment {
+
+    /**
+     * The experiment's sensitivity.
+     */
+    private double sensitivity;
+
+    /**
+     * The experiment's negative prediction rate.
+     */
+    private double npr;
+
+    /**
+     * A poisson distributed random number generator.
+     */
+    private Poisson poisson;
+
+    /**
+     * A uniform random number generator.
+     */
+    private Uniform uniform;
+
+    /**
+     * constructor
+     */
+    public ReductionistExperiment() {
+        poisson = new Poisson(250, RandomEngineRegistry.getEngine());
+        uniform = new Uniform(RandomEngineRegistry.getEngine());
+    }
+
+
+    /**
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/experiments/SimpleExperiment.java

@@ -1,0 +1,155 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import cern.jet.random.Uniform;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ * A basic implementation of the <code>Experiment</code> interface. The experiment
+ * introduces errors according to given sensitivity and specifity values.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Extension(id="simpleexperiment")
+public final class SimpleExperiment implements Experiment {
+
+
+    /**
+     * The experiment's sensitivity.
+     */
+    private double sensitivity;
+
+    /**
+     * The experiment's specificity.
+     */
+    private double specificity;
+
+    /**
+     * A random number generator.
+     */
+    private Uniform uniform = new Uniform(0.0, 1.0, RandomEngineRegistry.getEngine());
+
+
+    /**
+     * Gets the sensitivity.
+     * @return the sensitivity.
+     */
+    public double getSensitivity() {
+        return sensitivity;
+    }
+
+    /**
+     * Gets the specificity.
+     * @return the specificity.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/extsupport/Configurable.java

@@ -1,0 +1,34 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.extsupport;
+
+/**
+ * An extension to the harness which is configurable.
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public interface Configurable {
+
+    /**
+     * Returns a <code>String</code> containing a summary of the extension's
+     * configuration.
+     * @return a <code>String</code> containing a summary of the extension's
+     * configuration.
+     */
+    String configSummary();
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/extsupport/Extension.java

@@ -1,0 +1,41 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.extsupport;
+
+import java.lang.annotation.ElementType;
+import java.lang.annotation.Retention;
+import java.lang.annotation.RetentionPolicy;
+import java.lang.annotation.Target;
+
+/**
+ * Annotates a class as an an extension to the harness. Possible extensions should
+ * be subclasses of <code>Experiment</code>, <code>Graph</code>,
+ * <code>GraphPopulator</code> or <code>IntegrationMethod</code>. The
+ * <code>id</code> attribute is used to assign a short ID to the extension class.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Retention(RetentionPolicy.RUNTIME)
+@Target(ElementType.TYPE)
+public @interface Extension {
+    /**
+     * Returns the extension's ID.
+     * @return The extension's ID.
+     */
+    String id();
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/extsupport/ExtensionProperty.java

@@ -1,0 +1,43 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.extsupport;
+
+import java.lang.annotation.ElementType;
+import java.lang.annotation.Retention;
+import java.lang.annotation.RetentionPolicy;
+import java.lang.annotation.Target;
+
+/**
+ * This annotation is used to annotate setter methods that can be used to set the
+ * basic properties of an extension. For example, a method
+ * <code>setSensitivity(double sens)</code> could be annotated with
+ * <code>@ExtensionProperty(id="sensitivity")</code>.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Retention(RetentionPolicy.RUNTIME)
+@Target(ElementType.METHOD)
+public @interface ExtensionProperty {
+
+    /**
+     * The ID of the Extension property.
+     * @return The ID of the Extension property.
+     */
+    String id();
+    
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/extsupport/ExtensionRegistry.java

@@ -1,0 +1,400 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.extsupport;
+
+import java.io.IOException;
+import java.lang.reflect.Method;
+import java.util.HashMap;
+import java.util.Map;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import org.springframework.core.io.Resource;
+import org.springframework.core.io.support.PathMatchingResourcePatternResolver;
+import org.springframework.core.type.classreading.MetadataReader;
+import org.springframework.core.type.classreading.MetadataReaderFactory;
+import org.springframework.core.type.classreading.SimpleMetadataReaderFactory;
+import org.springframework.core.type.filter.AnnotationTypeFilter;
+import org.springframework.core.type.filter.AssignableTypeFilter;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.experiments.Experiment;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.populators.GraphPopulator;
+import uk.ac.ncl.cs.harness.integration.IntegrationMethod;
+import uk.ac.ncl.cs.harness.workflow.Properties;
+
+/**
+ * <p>The extension registry is a singleton object that can be queried for extensions
+ * to the harness. Extensions are subclasses of <code>Experiment</code>, <code>Graph</code>,
+ * <code>GraphPopulator</code> or <code>IntegrationMethod</code>. All extensions are
+ * registered with their ID (as defined by the <code>@Extension</code> annotation.</p>
+ * 
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/Edge.java

@@ -1,0 +1,167 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+import java.util.Collections;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Map;
+import java.util.Set;
+
+/**
+ * This class represents an edge in an undirected graph.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class Edge implements Comparable<Edge> {
+
+    /**
+     * The nodes that are connected by the edge.
+     */
+    private final SetOfTwo<Node> ij;
+
+    /**
+     * the probabilities carried by the edge.
+     */
+    private Map<String,Probability> probabilities;
+
+    /**
+     * the graph object to which this edge belongs.
+     */
+    private Graph owningGraph;
+
+    /**
+     * a package-private constructor that can only be used by graph implementations.
+     * The order of the connected nodes i and j does not matter, as this is an undirected edge.
+     *
+     * @param i a node
+     * @param j another node
+     * @param owningGraph the graph to which this edge (and the adjacent nodes) belong.
+     */
+    Edge(Node i, Node j, Graph owningGraph) {
+//        if (i.getId().compareTo(j.getId()) <= 0) {
+//            this.i = i;
+//            this.j = j;
+//        } else {
+//            this.i = j;
+//            this.j = i;
+//        }
+        ij = new SetOfTwo<Node>(i, j);
+        this.owningGraph = owningGraph;
+    }
+
+    /**
+     * Returns the two nodes, which are connected by this undirected edge.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/Graph.java

@@ -1,0 +1,115 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+import java.util.Collection;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * <p>This interface represents an undirected graph.</p>
+ * 
+ * <p>The Graph interface can be implemented as a possible extension to the
+ * harness. All implementing classes should carry the <code>@Extension</code> annotation
+ * as well as mark all their vital bean setters with <code>@ExtensionProperty</code>
+ * annotations.</p>
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public interface Graph {
+
+    /**
+     * sets the name of the graph.
+     * @param name the name
+     */
+    void setName(String name);
+
+    /**
+     * gets the name of the graph.
+     * @return the name
+     */
+    String getName();
+
+    /**
+     * creates a node with the given id
+     * @param id the node's id
+     * @return the node
+     */
+    Node createNode(String id);
+
+    /**
+     * Creates an edge between the given pair of nodes.
+     * @param nodes the pair of nodes which is to be connected by an edge.
+     * @return the edge
+     */
+    Edge createEdge(SetOfTwo<Node> nodes);
+
+    /**
+     * gets the set of all nodes contained in the graph.
+     * @return the set of all nodes contained in the graph.
+     */
+    Collection<Node> getNodes();
+
+    /**
+     * gets the set of all edges in the graph.
+     * @return the set of all edges in the graph.
+     */
+    Collection<Edge> getEdges();
+
+    /**
+     * gets the number of edges.
+     * @return the number of edges.
+     */
+    int getNumEdges();
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/MapGraph.java

@@ -1,0 +1,278 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+import java.util.*;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ * This class is a basic implementation of the <code>Graph</code> interface which
+ * stores all elements in maps. It is comparably inefficient for small graphs, but
+ * increasingly efficient for large, sparse graphs.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Extension(id="mapgraph")
+public final class MapGraph implements Graph {
+
+
+    /**
+     * The Graph's name
+     */
+    private String name;
+
+    /**
+     * Node id linked to node
+     */
+    private Map<String,Node> nodes = new HashMap<String, Node>();
+
+    /**
+     * Key (pair of node) linked to edge
+     */
+    private Map<SetOfTwo<Node>,Edge> edges = new HashMap<SetOfTwo<Node>, Edge>();
+
+    /**
+     * Node linked its neighbours
+     */
+    private Map<Node,Set<Neighbour>> neighbours = new HashMap<Node, Set<Neighbour>>();
+
+    /**
+     * gets the graph name
+     * @return
+     */
+    @Override
+    public String getName() {
+        return name;
+    }
+
+    /**
+     * sets the graph name
+     * @param name
+     */
+    @Override
+    public void setName(String name) {
+        this.name = name;
+    }
+
+    /**
+     * Creates a new node
+     * @param id
+     * @return
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/Neighbour.java

@@ -1,0 +1,102 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+/**
+ *
+ * This class models a neighbourhood in a graph. it stores an edge and the node
+ * that this edge connects to.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class Neighbour {
+
+    /**
+     * The connected node.
+     */
+    private Node node;
+
+    /**
+     * the connecting edge.
+     */
+    private Edge connectingEdge;
+
+    /**
+     * constructor with target node and connecting edge
+     * @param node the target node
+     * @param connectingEdge the connecting edge
+     */
+    Neighbour(Node node, Edge connectingEdge) {
+        this.node = node;
+        this.connectingEdge = connectingEdge;
+    }
+
+    /**
+     * returns the connecting edge.
+     * @return the connecting edge.
+     */
+    public Edge getConnectingEdge() {
+        return connectingEdge;
+    }
+
+    /**
+     * returns the target node.
+     * @return the target node.
+     */
+    public Node getNode() {
+        return node;
+    }
+
+    /**
+     * returns whether <code>obj</code> equals this neighbourhood.
+     * @param obj another object
+     * @return whether <code>obj</code> equals this neighbourhood.
+     */
+    @Override
+    public boolean equals(Object obj) {
+        if (obj == null) {
+            return false;
+        }
+        if (getClass() != obj.getClass()) {
+            return false;
+        }
+        final Neighbour other = (Neighbour) obj;
+        if (this.node != other.node && (this.node == null || !this.node.equals(other.node))) {
+            return false;
+        }
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/Node.java

@@ -1,0 +1,132 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+import java.util.Set;
+
+/**
+ * This class represents a node in a graph.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class Node implements Comparable<Node> {
+
+    /**
+     * the node's ID.
+     */
+    private final String id;
+
+    /**
+     * The graph to which this node belongs.
+     */
+    private final Graph owningGraph;
+
+    /**
+     * package-private constructor to be only used from graph implementations.
+     * @param id the node ID
+     * @param owningGraph the graph to which this node belongs.
+     */
+    Node(String id, Graph owningGraph) {
+        if(id == null) throw new NullPointerException("Can't create a node with null id");
+        if(owningGraph == null) throw new NullPointerException("Can't create a node with null graph");
+
+        this.id = id;
+        this.owningGraph = owningGraph;
+    }
+
+    /**
+     * gets the ID of this node
+     * @return the id of the node
+     */
+    public String getId() {
+        return id;
+    }
+
+    /**
+     * gets the set of all neighbours of this node in the network.
+     * @return the set of all neighbours of this node in the network
+     */
+    public Set<Neighbour> getNeighbours() {
+        return owningGraph.getNeighboursOfNode(this);
+    }
+
+    /**
+     * returns the node's degree (number of connecting edges)
+     * @return the node's degree
+     */
+    public int degree() {
+        return owningGraph.getNeighboursOfNode(this).size();
+    }
+
+    /**
+     * returns the graph to which this node belongs.
+     * @return the graph to which this node belongs.
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/Probability.java

@@ -1,0 +1,113 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+/**
+ * This class models a probability on a graph's edge
+ * @author Jochen Weile
+ */
+public class Probability {
+
+    /**
+     * Key for the edge's primary probability
+     */
+    public static final String MAIN_KEY = "MainProbability";
+
+    /**
+     * the probability's id.
+     */
+    private String key;
+
+    /**
+     * the actual probability value.
+     */
+    private double value;
+
+    /**
+     * constructor with id and probability value
+     * @param key
+     * @param value
+     */
+    public Probability(String key, double value) {
+        this.key = key;
+        this.value = value;
+    }
+
+    /**
+     * gets the key
+     * @return the key
+     */
+    public String getKey() {
+        return key;
+    }
+
+    /**
+     * sets the key
+     * @param key the key
+     */
+    public void setKey(String key) {
+        this.key = key;
+    }
+
+    /**
+     * sets the prob. value.
+     * @param value the prob. value.
+     */
+    public void setValue(double value) {
+        this.value = value;
+    }
+
+    /**
+     * gets the prob. value
+     * @return the prob. value.
+     */
+    public double getValue() {
+        return value;
+    }
+
+    /**
+     * equals method based on the key
+     */
+    @Override
+    public boolean equals(Object obj) {
+        if (obj == null) {
+            return false;
+        }
+        if (getClass() != obj.getClass()) {
+            return false;
+        }
+        final Probability other = (Probability) obj;
+        if ((this.key == null) ? (other.key != null) : !this.key.equals(other.key)) {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/populators/GraphPopulator.java

@@ -1,0 +1,49 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph.populators;
+
+import uk.ac.ncl.cs.harness.extsupport.Configurable;
+import uk.ac.ncl.cs.harness.graph.Graph;
+
+/**
+ * <p>Populators create nodes and edges in graphs according to specific rules.
+ * For example a possible populator could import network data from a file and use it to populate
+ * a graph object, or it could create nodes and edges on a random basis to form certain
+ * graph classifications, like random graphs or scale free graphs.</p>
+ *
+ * <p>The <code>GraphPopulator</code> interface can be implemented as a possible extension to the
+ * harness. All implementing classes should carry the <code>@Extension</code> annotation
+ * as well as mark all their vital bean setters with <code>@ExtensionProperty</code>
+ * annotations.</p>
+ *
+ * <p>To populate a graph with a populator, obtain an instance from the
+ * <code>ExtensionRegistry</code> and invoke its <code>populate(..)</code> method
+ * using the graph to fill as the parameter.</p>
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+
+public interface GraphPopulator extends Configurable  {
+
+    /**
+     * populates the given graph .
+     * @param graph the graph to populate.
+     */
+    void populate(Graph graph);
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/graph/populators/ScaleFreePopulator.java

@@ -1,0 +1,165 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph.populators;
+
+import cern.jet.random.Uniform;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.util.List;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ * This is a graph populator that generates a scale-free networks with the given
+ * number of seeds and a given number of nodes. the number of edges will equal the
+ * number of nodes, since this is an impementation of the preferential attachment
+ * algorithm.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Extension(id="scalefree")
+public final class ScaleFreePopulator implements GraphPopulator {
+
+//    /**
+//     * random number generator.
+//     */
+//    private Random random = new Random();
+    private Uniform uniform = new Uniform(RandomEngineRegistry.getEngine());
+
+    /**
+     * number of seed nodes to generate
+     */
+    private int seed;
+
+    /**
+     * number of nodes to generate
+     */
+    private int numberOfNodes;
+
+    /**
+     * gets the number of seed nodes.
+     * @return the number of seed nodes.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/EmpiricalIntegration.java

@@ -1,0 +1,91 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.Map;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+
+/**
+ * The empirical integration annotates each edge in the integrated graph with
+ * a probability equaling the proportion of times the edge is reported in the
+ * evidential graphs.
+ * 
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="empirical")
+public class EmpiricalIntegration extends IntegrationMethod {
+
+     /**
+     * Starts the integration process.
+     * @throws ConfigurationException if the integration method has been
+     * misconfigured.
+     */
+    @Override
+    public void start() throws ConfigurationException {
+
+        if (evidentialGraphs == null || evidentialGraphs.isEmpty()) {
+            throw new ConfigurationException("No evidental graphs given!");
+        }
+
+        performUnion();
+
+        double numEvGs = (double)evidentialGraphs.size();
+
+        for (Edge e : integratedGraph.getEdges()) {
+            double count = 0.0;
+            for (Graph evg : evidentialGraphs) {
+                if (evg.contains(e)) {
+                    count += 1.0;
+                }
+            }
+            double prob = count / numEvGs;
+            e.addProbability(Probability.MAIN_KEY, prob);
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/EmpiricalThresholdedIntegration.java

@@ -1,0 +1,89 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+
+/**
+ * The empirical thresholded integration annotates each edge in the integrated
+ * graph with 0 or 1 depending whether more or less than half the evidences
+ * vouch for it.
+ * 
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="empThr")
+public class EmpiricalThresholdedIntegration extends IntegrationMethod {
+
+     /**
+     * Starts the integration process.
+     * @throws ConfigurationException if the integration method has been
+     * misconfigured.
+     */
+    @Override
+    public void start() throws ConfigurationException {
+
+        if (evidentialGraphs == null || evidentialGraphs.isEmpty()) {
+            throw new ConfigurationException("No evidental graphs given!");
+        }
+
+        performUnion();
+
+        double numEvGs = (double)evidentialGraphs.size();
+
+        for (Edge e : integratedGraph.getEdges()) {
+            double count = 0.0;
+            for (Graph evg : evidentialGraphs) {
+                if (evg.contains(e)) {
+                    count += 1.0;
+                }
+            }
+            double prob = (double)count <= (double)numEvGs / 2.0 ? 0.0: 1.0;
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ImprovedLycettIntegration.java

@@ -1,0 +1,248 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.*;
+import uk.ac.ncl.cs.harness.stats.ErrorGauge;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ * This is an implementation of my amended version of Sam Lycett's
+ * probabilistic network integration algorithm. (Publication pending)
+ * 
+ * @author Jochen Weile, M.Sc.; (algorithm designed by Samantha Lycett, Ph.D.)
+ */
+@Extension(id="lycett2")
+public final class ImprovedLycettIntegration extends IntegrationMethod {
+
+    /**
+     * The logarithm of the prior odds that an edge exists in the network.
+     */
+    private double logPriorOdds;
+
+    /**
+     * Expected number of edges in the true network. Used to estimate the
+     * prior odds of edge existence.
+     */
+    private double expNumEdges;
+
+    private static final int L = 0;
+    private static final int NOT_L = 1;
+
+    /**
+     * Disables processing of negative evidence
+     */
+    private boolean disableNegativeEvidence = false;
+
+    /**
+     * starts the integration process.
+     * @throws ConfigurationException if the plugin has been misconfigured.
+     */
+    @Override
+    public void start() throws ConfigurationException {
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/IntegrationMethod.java

@@ -1,0 +1,153 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.extsupport.Configurable;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+
+/**
+ * <p>The <code>IntegrationMethod</code> interface can be implemented as a possible extension to the
+ * harness. All implementing classes should carry the <code>@Extension</code> annotation
+ * as well as mark all their vital bean setters with <code>@ExtensionProperty</code>
+ * annotations.</p>
+ *
+ * <p>Implementations of the integration methods usually take a set of evidential
+ * (optionally also a set of gold standard graphs) and try to infer the probabilities
+ * of edge existence in the union of these evidential graphs. To use an integration
+ * method, obtain an instance from the <code>ExtensionRegistry</code> supply it with
+ * the required evidential graphs, gold standards and other possible parameters. Also,
+ * you should provide an empty graph that will be filled to represent the integration result,
+ * using the method <code>setIntegratedGraph(..)</code>. Then execute the
+ * <code>start()</code> method.</p>
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public abstract class IntegrationMethod implements Configurable  {
+
+    /**
+     * an empty graph that will be filled with the integration results.
+     */
+    protected Graph integratedGraph;
+
+    /**
+     * the set of evidential graphs.
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/LeeIntegration.java

@@ -1,0 +1,242 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.Comparator;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+import uk.ac.ncl.cs.harness.stats.ErrorGauge;
+
+/**
+ * This is an implementation of the integration method described in
+ * <a href="http://www.sciencemag.org/content/306/5701/1555.abstract">
+ * Lee et al. 2006</a>.
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="lee")
+public class LeeIntegration extends IntegrationMethod {
+
+    /**
+     * The mysterious dependency weight value.
+     */
+    private double dValue;
+
+    /**
+     * Score quantification threshold.
+     */
+    private double threshold;
+
+    /**
+     * gets the dependency weight value.
+     * @return the dependency weight value.
+     */
+    public double getdValue() {
+        return dValue;
+    }
+
+    /**
+     * sets the dependency weight value.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/LycettIntegration.java

@@ -1,0 +1,184 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.graph.*;
+import uk.ac.ncl.cs.harness.stats.ErrorGauge;
+
+/**
+ * This is an implementation of Sam Lycett's probabilistic network integration
+ * algorithm. (Publication pending)
+ * 
+ * @author Jochen Weile, M.Sc.; (algorithm designed by Samantha Lycett, Ph.D.)
+ */
+@Extension(id="lycett")
+public final class LycettIntegration extends IntegrationMethod {
+
+    /**
+     * The logarithm of the prior odds that an edge exists in the network.
+     */
+    private static final double LOG_PRIOR_ODDS = 0.0;
+
+    /**
+     * starts the integration process.
+     * @throws ConfigurationException if the plugin has been misconfigured.
+     */
+    @Override
+    public void start() throws ConfigurationException {
+
+        if (evidentialGraphs == null || evidentialGraphs.isEmpty()) {
+            throw new ConfigurationException("No evidental graphs given!");
+        }
+
+        if (goldStandards == null || goldStandards.isEmpty()) {
+            throw new ConfigurationException("No gold standards given!");
+        }
+        
+        performUnion();
+
+        for (Graph goldStandard : goldStandards) {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/McmcFBM.java

@@ -1,0 +1,531 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import cern.jet.random.Beta;
+import cern.jet.random.Uniform;
+import java.io.IOException;
+
+import java.util.*;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.graph.Probability;
+import uk.ac.ncl.cs.harness.integration.ebm.BurnIn;
+import uk.ac.ncl.cs.harness.integration.ebm.EdgeAverager;
+import uk.ac.ncl.cs.harness.integration.ebm.ErrorRates;
+import uk.ac.ncl.cs.harness.integration.ebm.RateAverager;
+import uk.ac.ncl.cs.harness.io.OutputController;
+import uk.ac.ncl.cs.harness.stats.ErrorGauge;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * A Markov Chain Monte Carlo (MCMC)-based fully bayesian integration
+ * method.
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="mcmcEBM")
+public final class McmcFBM extends IntegrationMethod {
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/NaiveEBM.java

@@ -1,0 +1,179 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration;
+
+import java.util.Map;
+import java.util.HashMap;
+import java.util.Map.Entry;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+import uk.ac.ncl.cs.harness.integration.ebm.ErrorRates;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Extension(id="naiveEBM")
+public class NaiveEBM extends IntegrationMethod {
+
+    private Map<Graph,ErrorRates> errorRates;
+
+    private Map<Edge,Double> probabilities = new HashMap<Edge,Double>(1000);
+
+    @Override
+    public void start() {
+        performUnion();
+
+        initErrorRates();
+
+        double prior = 2.0 / ((double)integratedGraph.getNumNodes() - 1.0);
+
+        for (int i = 0; i < 1000; i++) {
+
+            updateProbabilities(prior);
+
+            updateErrorRates();
+
+            logErrorRates();
+        }
+
+        assignFinalValues();
+
+        invalidateRatesMap();
+
+    }
+
+//    private void generateTopologicalConsensus() {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ebm/BurnIn.java

@@ -1,0 +1,161 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration.ebm;
+
+import cern.colt.list.DoubleArrayList;
+import cern.jet.stat.Descriptive;
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Graph;
+
+/**
+ * Records the burn-in phase of the MCMC method and
+ * analyzes the autocorrelation to determine the best thinning factor.
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class BurnIn {
+
+    private static final int FPR_OFFSET = 0;
+    private static final int FNR_OFFSET = 1;
+
+    /**
+     * Determines the record indices associated with each graph.
+     */
+    private Map<Graph,Integer> graphIndices;
+
+    /**
+     * Matrix storing the error rates for each cycle.
+     * records[rateIndex][cycleIndex]
+     */
+    private double[][] records;
+
+    /**
+     * points to the next record. strictly internal use only.
+     */
+    private int recordPointer = 0;
+
+    /**
+     * the autocorrelation lag to use
+     */
+    private int lag;
+
+    /**
+     * Autocorrelation wrt given lag must be smaller than this threshold
+     * for all rates to accept current thinning rate.
+     */
+    private double acThreshold;
+
+    /**
+     * Constructs a new BurnIn instance.
+     * @param size number of burn-in cycles.
+     * @param numGraphs number of experimental graphs
+     * @param acLag autocorrelation lag to use
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ebm/EdgeAverager.java

@@ -1,0 +1,80 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration.ebm;
+
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Edge;
+
+/**
+ * Average over edge existence.
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class EdgeAverager {
+
+    /**
+     * Counts edges
+     */
+    private Map<Edge,Integer> edgeCounts = new HashMap<Edge, Integer>();
+
+    /**
+     * number of samples taken
+     */
+    private int numSamples = 0;
+
+    /**
+     * updates the counts
+     * @param edges set of currently existing edges
+     */
+    public void update(Set<Edge> edges) {
+        for (Edge e : edges) {
+            Integer count = edgeCounts.get(e);
+            if (count == null) {
+                count = 0;
+            }
+            edgeCounts.put(e,count+1);
+        }
+        numSamples++;
+    }
+
+    /**
+     * computes the averages
+     * @return a map linking each edge to its existence average
+     */
+    public Map<Edge,Double> getEdgeAverages() {
+
+        Map<Edge,Double> averages = new HashMap<Edge, Double>();
+
+        double numSamplesD = (double)numSamples;
+
+        for (Map.Entry<Edge,Integer> entry : edgeCounts.entrySet()) {
+            averages.put(entry.getKey(), (double)entry.getValue() / numSamplesD);
+        }
+
+        return averages;
+    }
+
+    public int getNumSamples() {
+        return numSamples;
+    }
+
+    
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ebm/ErrorRates.java

@@ -1,0 +1,78 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration.ebm;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ErrorRates {
+    private double fpr;
+    private double fnr;
+
+//    private double fprDensity, fnrDensity;
+
+//    public ErrorRates(double fpr, double fnr, double fprDensity, double fnrDensity) {
+//        this.fpr = fpr;
+//        this.fnr = fnr;
+//        this.fprDensity = fprDensity;
+//        this.fnrDensity = fnrDensity;
+//    }
+
+    public ErrorRates(double fpr, double fnr) {
+        this.fpr = fpr;
+        this.fnr = fnr;
+    }
+
+    public double getFalseNegativeRate() {
+        return fnr;
+    }
+
+    public double getFalsePositiveRate() {
+        return fpr;
+    }
+
+    private double lambdaNeg = Double.NaN, lambdaPos = Double.NaN;
+
+    public double getLogPositiveBayesFactor() {
+        if (Double.isNaN(lambdaPos)) {
+            lambdaPos = Math.log((1.0 - fnr) / fpr);
+        }
+        return lambdaPos;
+    }
+
+    public double getLogNegativeBayesFactor() {
+        if (Double.isNaN(lambdaNeg)) {
+            lambdaNeg = Math.log(fnr / (1.0 - fpr));
+        }
+        return lambdaNeg;
+    }
+
+//    public double getFalseNegativeRateDensity() {
+//        return fnrDensity;
+//    }
+//
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ebm/RandomWalkRecorder.java

@@ -1,0 +1,221 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration.ebm;
+
+import cern.colt.list.DoubleArrayList;
+import cern.jet.stat.Descriptive;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Graph;
+
+/**
+ * Records a random walk in the MCMC integrator
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+@Deprecated
+public class RandomWalkRecorder {
+
+    /**
+     * Assigns an ordinal number to each graph. Used internally to compute associated
+     * indices in storage arrays. (E.g if there are four graphs each of them gets a
+     * unique number between 0 and 3.
+     */
+    private Map<Graph,Integer> graphOrder = new HashMap<Graph, Integer>(8);
+
+    /**
+     * The data content. A list of arrays, each of which contains two fields for each graph
+     * plus one. Stores the false positive and false negative rates for each graph in each cycle.
+     * The last field is used to store the likelihood of teh cycle.
+     * E.g. given two graphs "0" and "1" each array contains
+     * [fp_0, fn_0, fp_1, fn_1, likelihood]
+     */
+    private List<double[]> data = new ArrayList<double[]>(1024);
+
+    private int skip = 1;
+
+    private int burnIn;
+
+    public RandomWalkRecorder(int burnIn) {
+        this.burnIn = burnIn;
+    }
+
+    
+
+//    /**
+//     * array storing the means for each value.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/integration/ebm/RateAverager.java

@@ -1,0 +1,75 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.integration.ebm;
+
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.stats.IncrementalAverage;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class RateAverager {
+
+    private static final int FPR_INDEX = 0;
+    private static final int FNR_INDEX = 1;
+
+    private IncrementalAverage[][] averages;
+
+    private Map<Graph,Integer> graph2Index = new HashMap<Graph, Integer>();
+
+    public RateAverager(Set<Graph> graphs) {
+        averages = new IncrementalAverage[2][graphs.size()];
+
+        int i = 0;
+        for (Graph g : graphs) {
+            graph2Index.put(g,i++);
+        }
+    }
+
+    public void update(Map<Graph,ErrorRates> rates) {
+        for (Graph graph : rates.keySet()) {
+            int graphIndex = graph2Index.get(graph);
+            double fpr = rates.get(graph).getFalsePositiveRate();
+            double fnr = rates.get(graph).getFalseNegativeRate();
+            __update(FPR_INDEX,graphIndex, fpr);
+            __update(FNR_INDEX,graphIndex, fnr);
+        }
+    }
+
+    private void __update(int rateIndex, int graphIndex, double d) {
+        if (averages[rateIndex][graphIndex] == null) {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/io/OutputController.java

@@ -1,0 +1,228 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.io;
+
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.FileOutputStream;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.OutputStream;
+import java.text.SimpleDateFormat;
+import java.util.Date;
+import java.util.HashMap;
+import java.util.Map;
+import java.util.logging.FileHandler;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+
+/**
+ * The OutputController is a central singleton that controls harness's output 
+ * behaviour. The singleton is obtained by calling the 
+ * <code>OutputController.getInstance();</code> method. 
+ * The <code>prepareOutput(String tag, File protocolFile)</code> can then be used
+ * to initialize a central output directory, which will contain the protocol input file,
+ * the workflow log file and all output files generated using the <code>writeResults</code>
+ * and <code>writeResultStream</code> methods.
+ *
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class OutputController {
+
+    /**
+     * current output directory.
+     */
+    private File currOutputDir;
+
+    /**
+     * singleton.
+     */
+    private static OutputController instance;
+
+    /**
+     * private constructor to enforce singleton.
+     */
+    private OutputController() {
+
+    }
+
+    /**
+     * singleton.
+     */
+    public static OutputController getInstance() {
+        if (instance == null) {
+            instance = new OutputController();
+        }
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/io/ProtocolParser.java

@@ -1,0 +1,353 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.io;
+
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.IOException;
+import java.io.InputStream;
+import java.util.Collections;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Iterator;
+import java.util.Map;
+import java.util.Set;
+import java.util.logging.Logger;
+import javax.xml.XMLConstants;
+import javax.xml.transform.Source;
+import javax.xml.transform.stream.StreamSource;
+import javax.xml.validation.Schema;
+import javax.xml.validation.SchemaFactory;
+import javax.xml.validation.Validator;
+import org.jdom.Document;
+import org.jdom.Element;
+import org.jdom.JDOMException;
+import org.jdom.Namespace;
+import org.jdom.input.SAXBuilder;
+import org.xml.sax.SAXException;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.workflow.Properties;
+import uk.ac.ncl.cs.harness.workflow.Protocol;
+import uk.ac.ncl.cs.harness.workflow.variables.IncrementalVariable;
+import uk.ac.ncl.cs.harness.workflow.variables.Type01LogarithmicVariable;
+import uk.ac.ncl.cs.harness.workflow.variables.Variable;
+
+/**
+ * Parses a given protocol XML file to create a <code>Protocol</code> 
+ * object.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class ProtocolParser {
+
+    /**
+     * a map linking variable identifiers to variables.
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/io/ResourcePaths.java

@@ -1,0 +1,35 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.io;
+
+import java.io.File;
+
+/**
+ * This class caches paths to the project's resource directories.
+ * 
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public interface ResourcePaths {
+
+    public static final File RESOURCES = new File(new File(new File("src"),"main"),"resources");
+
+    public static final File TEST_RESOURCES = new File(new File(new File("src"),"test"),"resources");
+
+    public static final File OUTPUT_DIR = new File("output");
+    
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/DegreeSample.java

@@ -1,0 +1,71 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.io.IOException;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.io.OutputController;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class DegreeSample {
+
+    private boolean done;
+
+    private DegreeSample() {
+
+    }
+
+    private static DegreeSample instance;
+
+    public static DegreeSample getInstance() {
+        if (instance == null) {
+            instance = new DegreeSample();
+        }
+        return instance;
+    }
+
+    public synchronized void sampleDegrees(Graph g) {
+        
+        if (done) {
+            return;
+        }
+
+        StringBuilder b = new StringBuilder();
+        for (Node node : g.getNodes()) {
+            int deg = node.getNeighbours().size();
+            b.append(deg).append("\n");
+        }
+
+        try {
+            OutputController.getInstance().writeResults("degrees.tsv", b.toString());
+        } catch (IOException ex) {
+            Logger.getLogger(DegreeSample.class.getName()).log(Level.SEVERE, "Unable to write degree results.", ex);
+        }
+
+        done = true;
+    }
+
+
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/ErrorGauge.java

@@ -1,0 +1,146 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.util.Collection;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ErrorGauge {
+
+    public static final int FP_RATE = 0, FN_RATE = 1,TP = 2,FP = 3,TN = 4,FN = 5;
+    
+    /**
+     *
+     * @param graphs
+     * @param trueGraph
+     * @return
+     */
+    public static Map<Graph,double[]> calculate(Set<Graph> graphs, Graph trueGraph) {
+
+        Map<Graph,double[]> ratesMap = new HashMap<Graph, double[]>();
+
+        for (Graph graph : graphs) {
+            double[] rates = calculate(graph,trueGraph);
+            ratesMap.put(graph, rates);
+        }
+
+        return ratesMap;
+    }
+
+
+    public static double[] calculate(Graph graph, Graph trueGraph) {
+        return calculate(graph,trueGraph.getEdges());
+    }
+
+    /**
+     * 
+     * @param graph
+     * @param trueGraph
+     * @return
+     */
+    public static double[] calculate(Graph graph, Collection<Edge> trueEdges) {
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/ErrorRateLossAverager.java

@@ -1,0 +1,66 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.util.Map;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import static uk.ac.ncl.cs.harness.stats.ErrorGauge.*;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ErrorRateLossAverager {
+
+    private IncrementalAverage average;
+
+    public void update(Map<Graph,double[]> trueRates, Map<Graph,double[]> expRates) {
+
+        double loss = 0.0;
+
+        for (Graph g : trueRates.keySet()) {
+
+            double fpr = trueRates.get(g)[FP_RATE];
+            double fnr = trueRates.get(g)[FN_RATE];
+
+            double fpr_meas = expRates.get(g)[FP_RATE];
+            double fnr_meas = expRates.get(g)[FN_RATE];
+
+            double fpr_diff = fpr - fpr_meas;
+            double fnr_diff = fnr - fnr_meas;
+
+            loss += fpr_diff * fpr_diff + fnr_diff * fnr_diff;
+
+        }
+
+        double n = trueRates.size() * 2;
+        loss = Math.sqrt(loss / n);
+
+        if (average == null) {
+            average = new IncrementalAverage(loss);
+        } else {
+            average.updateWith(loss);
+        }
+
+    }
+
+    public double getAverage() {
+        return average.getAverage();
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/FastROC.java

@@ -1,0 +1,83 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import cern.colt.list.DoubleArrayList;
+import java.awt.geom.Point2D;
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class FastROC {
+
+    public List<Point2D.Double> calculate(Graph integratedGraph, Graph trueGraph, double step) {
+
+        int v = trueGraph.getNumNodes();
+        int all = (v * v - v) / 2;
+        int baseNeg = all - integratedGraph.getNumEdges();
+
+        Set<Edge> set = new HashSet<Edge>(integratedGraph.getEdges());// O(|E|)
+        set.retainAll(trueGraph.getEdges());//set = tp  //O(|E|)
+        int baseFN = trueGraph.getNumEdges() - set.size();
+        int baseTN = baseNeg - baseFN;
+
+        DoubleArrayList reals = new DoubleArrayList();
+        DoubleArrayList nreals = new DoubleArrayList();
+
+        for (Edge e : integratedGraph.getEdges()) {//O(|E|)
+            double p = e.getProbability(Probability.MAIN_KEY).getValue();
+            if (trueGraph.contains(e)) {
+                reals.add(p);
+            } else {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/IncrementalAverage.java

@@ -1,0 +1,47 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class IncrementalAverage {
+
+    private double average;
+
+    private double i;
+
+    public IncrementalAverage(double initialValue) {
+        average = initialValue;
+        i = 1;
+    }
+
+    public void updateWith(double value) {
+        double n = (double)i;
+        average = n/(n+1) * average + value / (n+1);
+        i++;
+    }
+
+    public double getAverage() {
+        return average;
+    }
+
+    
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/RandomEngineRegistry.java

@@ -1,0 +1,52 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import cern.jet.random.engine.MersenneTwister;
+import cern.jet.random.engine.RandomEngine;
+import cern.jet.random.engine.RandomSeedGenerator;
+import java.util.HashMap;
+import java.util.Map;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class RandomEngineRegistry {
+
+    private static RandomSeedGenerator seeds = new RandomSeedGenerator();
+
+    private static Map<Thread,RandomEngine> engines = new HashMap<Thread, RandomEngine>(128);
+
+    public static RandomEngine getEngine() {
+        Thread t = Thread.currentThread();
+        RandomEngine engine = engines.get(t);
+        if (engine == null) {
+            engine = new MersenneTwister(seeds.nextSeed());
+            engines.put(t, engine);
+        }
+        return engine;
+    }
+
+    public static void deregister() {
+        Thread t = Thread.currentThread();
+        engines.remove(t);
+    }
+
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/ResultDistribution.java

@@ -1,0 +1,133 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.io.IOException;
+import java.util.HashSet;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+import uk.ac.ncl.cs.harness.io.OutputController;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ResultDistribution {
+
+    private int[] real;
+    private int[] nreal;
+
+    private static int slots;
+    public static synchronized void refreshSlots(int num) {
+        slots = num;
+    }
+    public static synchronized boolean acquireSlot() {
+        if (slots > 0) {
+            slots--;
+            return true;
+        } else {
+            return false;
+        }
+    }
+
+    public ResultDistribution(int bins) {
+        real = new int[bins+1];
+        nreal = new int[bins+1];
+    }
+
+    public void compute(Graph resultGraph, Graph trueGraph) {
+        double bins = real.length-1;
+
+        for (Edge e : resultGraph.getEdges()) {
+
+            double p = probabilityOf(e);
+            int index = (int)(bins * p);
+
+            if (trueGraph.contains(e)) {
+                real[index]++;
+            } else {
+                nreal[index]++;
+            }
+
+        }
+
+        Set<Edge> missing  = new HashSet<Edge>(trueGraph.getEdges());
+        missing.removeAll(resultGraph.getEdges());
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/ResultDistributionAverager.java

@@ -1,0 +1,71 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.io.IOException;
+import uk.ac.ncl.cs.harness.io.OutputController;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ResultDistributionAverager {
+
+    private IncrementalAverage[] real;
+    private IncrementalAverage[] nreal;
+
+    public ResultDistributionAverager(int numBins) {
+        real = new IncrementalAverage[numBins+1];
+        nreal = new IncrementalAverage[numBins+1];
+    }
+
+    public synchronized void update(ResultDistribution d) {
+        d.updateRealAverages(real);
+        d.updateNRealAverages(nreal);
+    }
+
+    public void writeOutput(String tag) {
+
+        double bins = real.length-1;
+
+        //compile output
+        StringBuilder b = new StringBuilder("p\treal\tnreal\n");
+        for (int i = 0; i < real.length; i++) {
+            double p = (double)i / bins;
+            b.append(p).append('\t')
+                    .append(real[i].getAverage()).append('\t')
+                    .append(nreal[i].getAverage()).append('\n');
+        }
+
+        //write output to file
+        String filename = "probDistr_"+tag+".tsv";
+        try {
+            OutputController.getInstance().writeResults(filename, b.toString());
+        } catch (IOException ex) {
+            throw new RuntimeException(
+                    "Unable to write probability distribution results to file: "
+                    +filename, ex);
+        }
+
+    }
+
+    public int getNumBins() {
+        return real.length-1;
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/stats/ResultLossAverager.java

@@ -1,0 +1,115 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import java.util.HashSet;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Probability;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ResultLossAverager {
+
+    private IncrementalAverage real;
+    private IncrementalAverage nreal;
+
+    public void update(Graph trueGraph, Graph integratedGraph) {
+
+        double realLoss = 0.0;
+        double nrealLoss = 0.0;
+
+        for (Edge e : integratedGraph.getEdges()) {
+            Probability pObj = e.getProbability(Probability.MAIN_KEY);
+            double p = pObj == null ? 0.0 : pObj.getValue();
+            if (trueGraph.contains(e)) {
+                double diff = 1.0 - p;
+                realLoss += diff * diff;
+            } else {
+                nrealLoss += p * p;
+            }
+        }
+
+        Set<Edge> fns = new HashSet<Edge>(trueGraph.getEdges());
+        fns.removeAll(integratedGraph.getEdges());
+        realLoss += fns.size();
+
+        int n = trueGraph.getNumNodes();
+        int all = ((n * n) - n) / 2;
+        int nrealAll = all - trueGraph.getNumEdges();
+
+        realLoss = Math.sqrt(realLoss / (double)trueGraph.getNumEdges());
+        nrealLoss = Math.sqrt(nrealLoss / (double)nrealAll);
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/util/IdService.java

@@ -1,0 +1,32 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.util;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class IdService {
+
+    private static long lastId = 0;
+
+    public static long getUniqueId() {
+        return ++lastId;
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/util/SetOfTwo.java

@@ -1,0 +1,112 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.util;
+
+/**
+ * A mathematical set of size two.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public final class SetOfTwo<T extends Comparable<T>> implements Comparable<SetOfTwo<T>> {
+
+    /**
+     * An element of the set.
+     */
+    private final T a, b;
+
+    /**
+     * constructor with to elements
+     * @param a
+     * @param b
+     */
+    public SetOfTwo(T a, T b) {
+        if(a == null || b == null) throw new NullPointerException("Can't construct a pair from null nodes");
+        if (a.compareTo(b) >= 0) {
+            this.a = a;
+            this.b = b;
+        } else {
+            this.a = b;
+            this.b = a;
+        }
+    }
+
+    /**
+     * gets an an element
+     * @return
+     */
+    public T getA() {
+        return a;
+    }
+
+    /**
+     * gets the other element
+     * @return
+     */
+    public T getB() {
+        return b;
+    }
+
+    /**
+     * returns whether this set equals another object
+     * @param obj whether this set equals another object
+     * @return
+     */
+    @Override
+    public boolean equals(Object obj) {
+        if (obj == null) {
+            return false;
+        }
+        if (getClass() != obj.getClass()) {
+            return false;
+        }
+        final SetOfTwo<T> other = (SetOfTwo<T>) obj;
+        if (!this.a.equals(other.a)) {
+            return false;
+        }
+        if (!this.b.equals(other.b)) {
+            return false;
+        }
+        return true;
+    }
+
+    /**
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/util/SynchronizedDoubleList.java

@@ -1,0 +1,39 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.util;
+
+import cern.colt.list.DoubleArrayList;
+
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class SynchronizedDoubleList {
+
+    private DoubleArrayList list = new DoubleArrayList();
+
+    public synchronized void add(double d) {
+        list.add(d);
+    }
+
+    public DoubleArrayList getArrayList() {
+        return list;
+    }
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/ConcurrentWorkflow.java

@@ -1,0 +1,220 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.io.IOException;
+import java.util.Collections;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.logging.Level;
+import java.util.Map;
+import java.util.concurrent.Semaphore;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.io.OutputController;
+import uk.ac.ncl.cs.harness.stats.ResultDistributionAverager;
+import uk.ac.ncl.cs.harness.stats.ErrorRateLossAverager;
+import uk.ac.ncl.cs.harness.stats.ResultLossAverager;
+import uk.ac.ncl.cs.harness.util.IdService;
+import uk.ac.ncl.cs.harness.workflow.variables.Variable;
+
+/**
+ * Executes a workflow protocol.
+ * @author Jochen Weile, M.Sc.
+ */
+public class ConcurrentWorkflow {
+
+    private int numCpuSlots;
+
+    private String[] variableIds;
+
+    private StringBuilder results = new StringBuilder();
+
+    /**
+     *
+     * @param numCpuSlots Number of CPUs to use at the same time.
+     */
+    public ConcurrentWorkflow(int numCpuSlots) {
+        this.numCpuSlots = numCpuSlots;
+    }
+
+    
+
+    /**
+     * runs a given workflow protocol
+     * @param protocol the protocol to execute.
+     * @throws ConfigurationException if the protocol is misconfigured.
+     */
+    public void run(Protocol protocol) throws ConfigurationException {
+
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/Properties.java

@@ -1,0 +1,123 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.util.HashMap;
+import java.util.Map;
+import java.util.Set;
+import uk.ac.ncl.cs.harness.workflow.variables.Variable;
+
+/**
+ * Stores a set of properties, which can potentially be mutable variables.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class Properties {
+
+    /**
+     * The key of to the ID property.
+     */
+    public static final String ID_PROPERTY = "ID_PROP";
+
+    /**
+     * The key to the "replicates" property.
+     */
+    public static final String REPLICAS_PROPERTY = "REPL_PROP";
+
+    /**
+     * internal map that stores the properties under their string keys, but
+     * references intermediate ObjRef objects, which resolve potential variables.
+     */
+    private Map<String,ObjRef> map = new HashMap<String, ObjRef>();
+
+    /**
+     * sets a property, this can either be an actual value or a variable.
+     * @param key
+     * @param val
+     */
+    public void setProperty(String key, Object val) {
+        map.put(key, new ObjRef(val));
+    }
+
+    /**
+     * gets a property
+     * @param key the property key
+     * @return the property value (either a static value or the current state
+     * of the referenced variable.
+     */
+    public Object getProperty(String key) {
+        return map.get(key).getValue();
+    }
+
+    /**
+     * gets the set of all stored property keys.
+     * @return
+     */
+    public Set<String> getKeys() {
+        return map.keySet();
+    }
+
+    /**
+     * internal helper class that provides a black box which abstracts over static
+     * values or variable references.
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/Protocol.java

@@ -1,0 +1,232 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.util.Map;
+import java.util.Set;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.workflow.variables.Variable;
+
+/**
+ * The protocol describes the details of the workflow execution. It defines variables,
+ * graph implementations and populators to use, experiments and integrations to perform.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public class Protocol {
+
+    /**
+     * the logger.
+     */
+    private static final Logger logger =
+            Logger.getLogger(Protocol.class.getName());
+
+    /**
+     * properties describing which graph populator to use and how to
+     * configure it.
+     */
+    private Properties graphPopulation;
+
+    /**
+     * id of the graph implementation to use.
+     */
+    private String graphImplementation;
+
+    /**
+     * properties describing which integration method to use and how to
+     * configure it.
+     */
+    private Properties integrationType;
+
+    /**
+     * set of properties describing which experiment configurations to use to
+     * simulate gold standards.
+     */
+    private Set<Properties> goldStandards;
+
+    /**
+     * set of properties describing which experiment configurations to use to
+     * simulate evidential graphs.
+     */
+    private Set<Properties> experiments;
+
+    /**
+     * A map linking variable IDs to the actual variables.
+     */
+    private Map<String, Variable> variables;
+
+    /**
+     * number of cycle replicas
+     */
+    private int cycleReplicas;
+
+
+    /**
+     * gets the graph implementation id
+     * @return the graph implementation id
+     */
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/ResultCallback.java

@@ -1,0 +1,51 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.util.HashMap;
+import uk.ac.ncl.cs.harness.stats.ErrorRateLossAverager;
+import uk.ac.ncl.cs.harness.stats.ResultDistributionAverager;
+import uk.ac.ncl.cs.harness.stats.ResultLossAverager;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ResultCallback extends HashMap<String,Object> {
+
+    private ResultLossAverager resultLossAverager = new ResultLossAverager();
+    private ResultDistributionAverager resultDistributionAverager = new ResultDistributionAverager(100);
+    private ErrorRateLossAverager rateLossAverager = new ErrorRateLossAverager();
+
+    public ErrorRateLossAverager getRateLossAverager() {
+        return rateLossAverager;
+    }
+
+    public ResultDistributionAverager getResultDistributionAverager() {
+        return resultDistributionAverager;
+    }
+
+    public ResultLossAverager getResultLossAverager() {
+        return resultLossAverager;
+    }
+
+
+    
+
+
+}

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/WorkflowThread.java

@@ -1,0 +1,268 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.util.Collections;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
+import java.util.concurrent.Semaphore;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.experiments.Experiment;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionRegistry;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.populators.GraphPopulator;
+import uk.ac.ncl.cs.harness.integration.IntegrationMethod;
+import uk.ac.ncl.cs.harness.io.OutputController;
+import uk.ac.ncl.cs.harness.stats.ResultDistributionAverager;
+import uk.ac.ncl.cs.harness.stats.DegreeSample;
+import uk.ac.ncl.cs.harness.stats.ErrorGauge;
+import uk.ac.ncl.cs.harness.stats.ErrorRateLossAverager;
+import uk.ac.ncl.cs.harness.stats.ResultLossAverager;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.stats.ResultDistribution;
+
+/**
+ * This thread executes a single workflow according to a given protocol.
+ * It is initiated with a protocol instance, a semaphore (used to signal
+ * the thread's completion), a list of exceptions, which is used to 
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/variables/IncrementalVariable.java

@@ -1,0 +1,107 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow.variables;
+
+/**
+ * This is a variable that starts with a minimum value and performs increment
+ * operations by a certain value at each step until a maximum value is reached.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public final class IncrementalVariable extends Variable<Double> {
+
+    /**
+     * the start value (minimal value)
+     */
+    double min;
+
+    /**
+     * The end value (maximum value)
+     */
+    double max;
+
+    /**
+     * the incremental value.
+     */
+    double inc;
+
+    /**
+     * the current value.
+     */
+    double currVal;
+
+    /**
+     * constructor
+     * @param id the id of the variable
+     * @param min the start value (minimal value)
+     * @param max The end value (maximum value)
+     * @param inc the incremental value.
+     */
+    public IncrementalVariable(String id, double min, double max, double inc) {
+        super(id);
+        this.min = min;
+        this.max = max;
+        this.inc = inc;
+
+        reset();
+    }
+
+    /**
+     * resets the variable to the start value.
+     */
+    @Override
+    public void reset() {
+        currVal = min;
+    }
+
+    /**
+     * whether or not the variable can perform another step operation.
+     * @return whether or not the variable can perform another step operation.
+     */
+    @Override
+    public boolean canStep() {
+        return currVal+inc <= max;
+    }
+
+    /**
+     * perform a step
+     */
+    @Override
+    public void step() {
+        if (canStep()) {
+            currVal += inc;
+        }
+    }
+
+    /**
+     * returns the variable's current value.
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/variables/Type01LogarithmicVariable.java

@@ -1,0 +1,98 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow.variables;
+
+/**
+ * This variable logarithmically approaches 1,
+ * taking the values 0.9, 0.99, 0.999 etc.
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public final class Type01LogarithmicVariable extends Variable<Double> {
+
+    /**
+     * The variable's TAG, used by the protocol parser
+     */
+    public static final String TAG = "0-1";
+
+    /**
+     * number of iterations to perform
+     */
+    private int iterations;
+
+    /**
+     * current iteration
+     */
+    private int currIt;
+
+    /**
+     * constructor with id and number iterations to perform.
+     * @param id the variable's ID.
+     * @param iterations number of iterations to perform.
+     */
+    public Type01LogarithmicVariable(String id, int iterations) {
+        super(id);
+        this.iterations = iterations;
+
+        reset();
+    }
+
+    /**
+     * resets to start value
+     */
+    @Override
+    public void reset() {
+        currIt = 1;
+    }
+
+    /**
+     * whether or not the variable can perform another step operation.
+     * @return whether or not the variable can perform another step operation.
+     */
+    @Override
+    public boolean canStep() {
+        return currIt <= iterations;
+    }
+
+    /**
+     * perform a step
+     */
+    @Override
+    public void step() {
+        if (canStep()) {
+            currIt++;
+        }
+    }
+
+    /**
+     * returns the variable's current value.
+     * @return the variable's current value.
+     */
+    @Override
+    public Double getCurrentValue() {
[…]

### 2011-08-19 src/main/java/uk/ac/ncl/cs/harness/workflow/variables/Variable.java

@@ -1,0 +1,79 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow.variables;
+
+/**
+ * <p>Variables can iterate over a number of values. Example usage:</p>
+ * <pre><code>
+ * Variable v = new Type01LogarithmicVariable("test",4);
+ * while (v.canStep() {
+ *    double val = v.getCurrentValue();
+ *    step();
+ * }
+ * </code></pre>
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+public abstract class Variable<T> {
+
+    /**
+     * the variable's ID
+     */
+    private String id;
+
+    /**
+     * Creates a variable with ID id
+     */
+    protected Variable(String id) {
+        this.id = id;
+    }
+
+    /**
+     * gets the variable's ID
+     */
+    public String getId() {
+        return id;
+    }
+
+    /**
+     * resets to start value
+     */
+    public abstract void reset();
+
+    /**
+     * whether or not the variable can perform another step operation.
+     */
+    public abstract boolean canStep();
+
+    /**
+     * step to the next value if possible
+     */
+    public abstract void step();
+
+    /**
+     * returns the variable's current value.
+     */
+    public abstract T getCurrentValue();
+
+    /**
+     * returns the number of total iterations through which this variable cycles.
+     */
+    public abstract int getNumTotalIterations();
+
+
+}

### 2011-08-19 src/main/r/plot.R

@@ -1,0 +1,20 @@
+#!/usr/bin/Rscript
+
+pdf("plot.pdf",width=4,height=8)
+op=par(mfrow=c(3,1))
+
+fn <- read.table("fn.tsv",header=TRUE,row.names=1,sep="\t")
+barplot(as.matrix(t(fn)),beside=TRUE,xlab="# experiments",ylab=expression(E(L^{"("+")"})) )
+legend("topright",c(expression(paste("Lee ",italic("et al."))),expression(paste("James ",italic("et al."))),expression(paste("Weile ",italic("et al.")))),bty="n",fill=grey.colors(3))
+
+fp <- read.table("fp.tsv",header=TRUE,row.names=1,sep="\t")
+barplot(as.matrix(t(fp)),beside=TRUE,xlab="# experiments",ylab=expression(E(L^{"("-")"})) )
+#legend("right",colnames(fp),bty="n",fill=grey.colors(3))
+
+er <- read.table("erloss.tsv",header=TRUE,row.names=1,sep="\t")
+barplot(as.matrix(t(er)),beside=TRUE,xlab="# experiments",ylab=expression(E(L["ER"])) )
+#legend("right",colnames(er),bty="n",fill=grey.colors(3))
+
+par(op)
+dev.off()
+

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/HarnessIT.java

@@ -1,0 +1,77 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness;
+
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.InputStream;
+import java.io.InputStreamReader;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.io.ResourcePaths;
+
+/**
+ *
+ * @author jweile
+ */
+public class HarnessIT extends TestCase {
+    
+    public HarnessIT(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testIntegration() throws Exception {
+
+        File jarFile = new File(new File("target"),
+                "harness-0.0.1-SNAPSHOT-jar-with-dependencies.jar");
+
+        File protocolFile = new File(ResourcePaths.TEST_RESOURCES, "protocol.xml");
+
+        String[] execArgs = new String[] {"java", "-jar", 
+                                jarFile.getCanonicalPath(), "-t failsafe",
+                                protocolFile.getCanonicalPath()};
+
+        Process process = Runtime.getRuntime().exec(execArgs);
+
+        InputStream errorStream = process.getErrorStream();
+        BufferedReader errorReader = new BufferedReader(
+                                        new InputStreamReader(errorStream)
+                                     );
+
+        String line;
+        while ((line = errorReader.readLine()) != null) {
+            System.err.println(line);
+        }
+        errorReader.close();
+
+        process.waitFor();
+
+        assertEquals(0, process.exitValue());
+
+    }
+
+}

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/experiments/FastExperimentTest.java

@@ -1,0 +1,120 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import java.util.HashSet;
+import java.util.Set;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+import uk.ac.ncl.cs.harness.graph.populators.ScaleFreePopulator;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class FastExperimentTest extends TestCase {
+
+    private Graph graph;
+
+    public FastExperimentTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+
+        super.setUp();
+
+        graph = new MapGraph();
+        graph.setName("main");
+        ScaleFreePopulator populator = new ScaleFreePopulator();
+        populator.setSeed(3);
+        populator.setNumberOfNodes(500);
+        populator.populate(graph);
+
+        Logger.getLogger(this.getClass().getName())
+                .log(Level.INFO, "Template edges: {0}",
+                     graph.getNumEdges());
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+        graph = null;
+    }
+
+    /**
+     * Test of perform method, of class SimpleExperiment.
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/experiments/ReductionistExperimentTest.java

@@ -1,0 +1,128 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import java.util.HashSet;
+import java.util.Set;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+import uk.ac.ncl.cs.harness.graph.populators.ScaleFreePopulator;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ReductionistExperimentTest extends TestCase {
+
+    /**
+     * the graph
+     */
+    private Graph graph;
+
+    public ReductionistExperimentTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+
+        super.setUp();
+
+        graph = new MapGraph();
+        graph.setName("main");
+        ScaleFreePopulator populator = new ScaleFreePopulator();
+        populator.setSeed(3);
+        populator.setNumberOfNodes(500);
+        populator.populate(graph);
+
+        Logger.getLogger(this.getClass().getName())
+                .log(Level.INFO, "Template edges: {0}",
+                     graph.getNumEdges());
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+        graph = null;
+    }
+
+    /**
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/experiments/SimpleExperimentTest.java

@@ -1,0 +1,77 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+import uk.ac.ncl.cs.harness.graph.populators.ScaleFreePopulator;
+
+/**
+ *
+ * @author jweile
+ */
+public class SimpleExperimentTest extends TestCase {
+
+    private Graph graph;
+
+    public SimpleExperimentTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+
+        LoggingInit.init();
+
+        super.setUp();
+
+        graph = new MapGraph();
+        graph.setName("main");
+        ScaleFreePopulator populator = new ScaleFreePopulator();
+        populator.setSeed(3);
+        populator.setNumberOfNodes(500);
+        populator.populate(graph);
+
+        System.out.println("Template edges: "+graph.getNumEdges());
+
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+
+        graph = null;
+    }
+
+    /**
+     * Test of perform method, of class SimpleExperiment.
+     */
+    public void testPerform() {
+        SimpleExperiment experiment = new SimpleExperiment();
+        experiment.setSensitivity(.99);
+        experiment.setSpecificity(.9997);
+        Graph experimentalGraph = new MapGraph();
+        experimentalGraph.setName("experimental");
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/extsupport/ExtensionRegistryTest.java

@@ -1,0 +1,51 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.extsupport;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.graph.Graph;
+
+/**
+ *
+ * @author jweile
+ */
+public class ExtensionRegistryTest extends TestCase {
+    
+    public ExtensionRegistryTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testRegistry() throws ConfigurationException {
+        Graph g = ExtensionRegistry.getInstance().getGraphImplementation("mapgraph");
+        g.createNode("test");
+    }
+
+}

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/graph/MapGraphTest.java

@@ -1,0 +1,116 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * @author jweile
+ */
+public class MapGraphTest extends TestCase {
+    
+    public MapGraphTest(String testName) {
+        super(testName);
+    }
+
+    private Graph graph;
+
+    @Override
+    protected void setUp() throws Exception {
+
+        LoggingInit.init();
+
+        super.setUp();
+        
+        graph = new MapGraph();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    /**
+     * Test of setName method, of class MapGraph.
+     */
+    public void testName() {
+        String name = "test";
+        graph.setName(name);
+        assertEquals(name, graph.getName());
+    }
+
+
+    /**
+     * Test of createNode method, of class MapGraph.
+     */
+    public void testNodesAndEdges() {
+        Node a = graph.createNode("A");
+        Node b = graph.createNode("B");
+
+        assertEquals(2,graph.getNumNodes());
+
+        SetOfTwo<Node> ab = new SetOfTwo<Node>(a,b);
+        Edge e = graph.createEdge(ab);
+
+        assertTrue(graph.contains(a));
+        assertTrue(graph.contains(b));
+        assertTrue(graph.contains(e));
+
+        assertEquals(1, graph.getNumEdges());
+
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/graph/populators/ScaleFreePopulatorTest.java

@@ -1,0 +1,74 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.graph.populators;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+
+/**
+ *
+ * @author jweile
+ */
+public class ScaleFreePopulatorTest extends TestCase {
+
+    private GraphPopulator pop;
+
+    public ScaleFreePopulatorTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+
+    /**
+     * Test of populate method, of class ScaleFreePopulator.
+     */
+    public void testPopulate() {
+
+        int seed = 4;
+        int numNodes = 500;
+
+
+        Graph graph = new MapGraph();
+        graph.setName("test");
+
+        pop = new ScaleFreePopulator();
+        ((ScaleFreePopulator)pop).setSeed(seed);
+        ((ScaleFreePopulator)pop).setNumberOfNodes(numNodes);
+
+        pop.populate(graph);
+        
+        assertEquals("Number of nodes", numNodes, graph.getNumNodes());
+
+        int surplusEdges = (((seed * seed)- seed) / 2) - seed;
+        assertEquals("Number of edges", numNodes+surplusEdges, graph.getNumEdges());
+
+
+    }
+
+
+}

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/io/ProtocolParserTest.java

@@ -1,0 +1,99 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.io;
+
+import java.io.File;
+import java.util.logging.Logger;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.exceptions.ConfigurationException;
+import uk.ac.ncl.cs.harness.workflow.Properties;
+import uk.ac.ncl.cs.harness.workflow.Protocol;
+
+/**
+ *
+ * @author jweile
+ */
+public class ProtocolParserTest extends TestCase {
+    
+    public ProtocolParserTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testValidation() throws Exception {
+
+        File protocolFile = new File(ResourcePaths.TEST_RESOURCES,
+                "protocol.xml");
+
+        ProtocolParser parser = new ProtocolParser();
+
+        parser.validate(protocolFile);
+
+        protocolFile = new File(ResourcePaths.TEST_RESOURCES,
+                "invalid_protocol.xml");
+
+        ConfigurationException conf = null;
+        try {
+            parser.validate(protocolFile);
+        } catch(ConfigurationException e) {
+            conf = e;
+            Logger.getLogger(this.getClass().getName()).info(e.getMessage());
+        }
+        assertNotNull(conf);
+
+    }
+
+    public void testParsing() throws Exception {
+
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/stats/ErrorGaugeTest.java

@@ -1,0 +1,87 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+import static uk.ac.ncl.cs.harness.stats.ErrorGauge.*;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class ErrorGaugeTest extends TestCase {
+    
+    public ErrorGaugeTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    
+    public void testErrorRates() {
+
+        Graph trueGraph = new MapGraph();
+
+        Node n1 = trueGraph.createNode("1");
+        Node n2 = trueGraph.createNode("2");
+        Node n3 = trueGraph.createNode("3");
+        Node n4 = trueGraph.createNode("4");
+
+        trueGraph.createEdge(new SetOfTwo<Node>(n1,n2));
+        trueGraph.createEdge(new SetOfTwo<Node>(n2,n4));
+        trueGraph.createEdge(new SetOfTwo<Node>(n1,n3));
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/stats/IncrementalAverageTest.java

@@ -1,0 +1,72 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.stats;
+
+import junit.framework.TestCase;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class IncrementalAverageTest extends TestCase {
+    
+    public IncrementalAverageTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+     public void testAverage() {
+         IncrementalAverage av = new IncrementalAverage(1.0);
+
+         av.updateWith(1.0);
+         assertAlmostEqual(1.0, av.getAverage());
+
+         av.updateWith(1.0);
+         assertAlmostEqual(1.0, av.getAverage());
+
+         av.updateWith(0.0);
+         assertAlmostEqual(3.0 / 4.0, av.getAverage());
+
+         av.updateWith(0.0);
+         assertAlmostEqual(3.0 / 5.0, av.getAverage());
+
+         av.updateWith(1.0);
+         assertAlmostEqual(4.0 / 6.0, av.getAverage());
+
+         av.updateWith(0.0);
+         assertAlmostEqual(4.0 / 7.0, av.getAverage());
+     }
+
+     private void assertAlmostEqual(double expected, double found) {
+         double diff = Math.abs(expected - found);
+         double maxDiff = .000000000000001;
+         if (diff > maxDiff) {
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/workflow/ConcurrentWorkflowTest.java

@@ -1,0 +1,111 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import java.util.Collections;
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.Map;
+import java.util.Set;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.io.OutputController;
+import uk.ac.ncl.cs.harness.workflow.variables.IncrementalVariable;
+import uk.ac.ncl.cs.harness.workflow.variables.Variable;
+
+/**
+ *
+ * @author jweile
+ */
+public class ConcurrentWorkflowTest extends TestCase {
+    
+    public ConcurrentWorkflowTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+        OutputController.getInstance().prepareOutput("surefire", null);
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testStaticWorkflow() throws Exception {
+
+        Protocol protocol = makeProtocol();
+
+        ConcurrentWorkflow workflow = new ConcurrentWorkflow(4);
+        workflow.run(protocol);
+
+    }
+
+    public void testDynamicWorkflow() throws Exception {
+
+        String varName = "var";
+        Map<String,Variable> variables = new HashMap<String, Variable>();
+        Variable var = new IncrementalVariable(varName, 2.0, 3.0, 1.0);
+        variables.put(varName, var);
+
[…]

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/workflow/PropertiesTest.java

@@ -1,0 +1,63 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow;
+
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.workflow.variables.IncrementalVariable;
+
+/**
+ *
+ * @author jweile
+ */
+public class PropertiesTest extends TestCase {
+    
+    public PropertiesTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testProperties() {
+        
+        Properties properties = new Properties();
+        
+        IncrementalVariable var2 = new IncrementalVariable("2", 2.0, 3.0, 1.0);
+        
+        properties.setProperty("1", 1);
+        properties.setProperty("2", var2);
+        
+        assertEquals(1, properties.getProperty("1"));
+        assertEquals(2.0, properties.getProperty("2"));
+        
+        var2.step();
+        
+        assertEquals(3.0, properties.getProperty("2"));
+
+    }
+
+}

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/workflow/variables/IncrementalVariableTest.java

@@ -1,0 +1,60 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow.variables;
+
+import junit.framework.TestCase;
+
+/**
+ *
+ * @author jweile
+ */
+public class IncrementalVariableTest extends TestCase {
+    
+    public IncrementalVariableTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testVariable() {
+        double min = 1.0, max = 5.0, inc = 1.0;
+
+        Variable<Double> var = new IncrementalVariable("test", min, max, inc);
+
+        int i = 1;
+        double currVal = min;
+        while (var.canStep()) {
+
+            assertEquals(currVal, var.getCurrentValue());
+            assertTrue(var.getCurrentValue() <= max);
+
+            var.step(); currVal += inc; i++;
+        }
+        assertEquals(currVal, 5.0);
+        assertEquals(var.getNumTotalIterations(), i);
+    }
+
+}

### 2011-08-19 src/test/java/uk/ac/ncl/cs/harness/workflow/variables/Type01LogarithmicVariableTest.java

@@ -1,0 +1,55 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.workflow.variables;
+
+import junit.framework.TestCase;
+
+/**
+ *
+ * @author jweile
+ */
+public class Type01LogarithmicVariableTest extends TestCase {
+    
+    public Type01LogarithmicVariableTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    public void testVariable() {
+        int iterations = 5;
+        Variable<Double> var = new Type01LogarithmicVariable("test", iterations);
+
+        int i = 0;
+        while (var.canStep()) {
+            System.out.println(var.getCurrentValue());
+            var.step();
+            i++;
+        }
+        assertEquals(iterations, i);
+    }
+
+}

### 2011-08-19 src/main/latex/paper/paper.tex
Incorporating Darren's changes in the paper

@@
 We present a novel algorithm for computing the probability of experimental observations to be correct that operates without gold standard reference data.
-We show that our algorithm even outperforms existing gold standard based methods.
+We show that our algorithm outperforms existing gold standard based methods.
 Finally, we apply the new algorithm on a large collection of genetic interaction and protein-protein interaction experiments.
@@
 
-A significant portion of knowledge about molecular biological processes is distributed over a large number of online databases \citep{stein_creating_2002}.
+A significant proportion of knowledge about molecular biological processes is distributed over a large number of online databases \citep{stein_creating_2002}.
 This knowledge has been obtained through experiments performed in laboratories all over the world.
@@
 The sub-discipline of integrative bioinformatics aims at collating this knowledge and making it accessible to both humans and computers.
-A large number of approaches to this problem exists.
-%\note{Mention some}.
 
 A popular integration paradigm is the construction of functional networks \citep{von_mering_string:_2003,warde-farley_genemania_2010}\note{Also cite Katherine}.
-Functional networks represent several different types of relationships between biological entities in an abstract manner.
-Associations such as genetic interactions, protein-protein interactions, gene regulation and co-expression, are combined into abstract statements of functional relatedness, which are termed functional interactions.
+Functional networks represent different types of relationships between biological entities in an abstract manner.
+Associations such as genetic interactions, protein-protein interactions, gene regulation and co-expression are combined into simple abstract statements of functional relatedness, which are termed functional interactions.
 
@@
 
-To provide a more generic solution to this problem, we present a fully Bayesian method that calculates, for each statement in a semantically-integrated dataset, the probability that it is true.
+To provide a more generic solution to this problem, we present a fully Bayesian method which calculates, for each statement in a semantically-integrated dataset, the probability that it is true.
 We have evaluated the method's effectiveness in comparison to related methods.
-The validity of any results of the method's application to real data is hard to verify without knowing the absolute biological truth.
+The validity of any results of the method's application to real data is difficult to verify without knowing the absolute biological truth.
 Therefore, we have developed a tool that tests integration methods on simulated data with the same characteristics as real biological networks.
@@
 If one considers each pair of entities to potentially have an edge, it is possible to model the process of prediction of such an edge by an experiment as follows.
-Let $X = \{X_1, \ldots, X_n \}$ be a vector of $n$ networks, which have been experimentally derived from $G$.
+Let $X = \{X_1, \ldots, X_n \}$ be a collection of $n$ networks which have been experimentally derived from $G$.
 Considering a single potential edge $e$, each experiment $X_i$ makes a statement about $e$'s existence.
@@
 
-An important concept necessary to determine this probability is the Bayes factor \citep{goodman_toward_1999}.
+An important concept necessary to determine this probability is the Bayes factor \citep{kass_bayes_1995}.
 For each of the $n$ experiments a Bayes factor $\Lambda_i$ can be determined, which is defined as follows:
@@
 
-However, instead of sampling from $\pi(\theta,G|X)$ directly, we employ Gibbs sampling \citep{hastings_monte_1970} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.
+However, instead of sampling from $\pi(\theta,G|X)$ directly, we employ Gibbs sampling \citep{gelfand_sampling-based_1990} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.
 
@@
 In order to obtain values for $a$ and $b$, one can compare each $X_i$ to the currently assumed true graph $G$ and use it to count the number of supposed True Positives (tp), False Positives (fp), True Negatives (tn) and False Negatives (fn).
-Then, $\alpha$ and $\beta$ can be sampled as follows:
+Then, the full conditionals for $\alpha$ and $\beta$ can be sampled as follows:
 \begin{align}
@@
 To initiate the algorithm as a whole, we need to generate initial error rate vector $\theta_{\text{init}}$.
-It is sufficient to sample the initial values for each $\alpha_i$ and $\beta_i$ from $Be(1,1)$, which is equivalent to the uniform distribution $\mathcal{U}_{[0,1]}$.
+It is sufficient to sample the initial values for each $\alpha_i$ and $\beta_i$ from their prior distribution.
+%$Be(1,1)$, which is equivalent to the uniform distribution $\mathcal{U}_{[0,1]}$. 
 Pseudocode for the complete algorithm is given in the supplementary material.
@@
 In order to evaluate the method described above and to compare it against other probabilistic integration methods, a simulation and testing environment was created.
-The testing tool creates a random graph according to a specified topology.
+The testing tool creates a random graph according to a specified model.
 In the simulation, this graph assumes the role of a true biological graph.
@@
 
-The simulated true graph is created as a scale-free graph using a preferential attachment algorithm, since most molecular-biological graphs have been shown to be belong to the scale-free class \citep{jeong_lethality_2001, eisenberg_preferential_2003}.
+The simulated true graph is created as a scale-free graph using a preferential attachment algorithm, since most molecular-biological graphs have been shown to be approximately scale-free \citep{jeong_lethality_2001, eisenberg_preferential_2003}.
 A description of the algorithm can be found in the supplementary material.
@@
 There are a number of different quality measures that can be applied.
-One important aspect to measure the accuracy of the method to estimate the error rates in the individual experiments.
-We can define the quadratic loss regarding the error estimates as follows.
+One important aspect is to measure the accuracy of the method in estimating the error rates in the individual experiments.
+We can define the quadratic loss for the error estimates as follows:
 \begin{equation}
   L_\text{ER} = \frac{1}{2n}\sum_{i=1}^n (\alpha_i - \hat{\alpha_i})^2 + (\beta_i - \hat{\beta_i})^2
-\end{equation}
+\end{equation},
 where $\hat{\alpha_i}$ and $\hat{\beta_i}$ are the estimates of the false positive and false negative rates for experiment $i$ according to the integration method in question.
@@
 Since we can expect a vast number of true negatives when working with sparse, scale-free graphs, it would be helpful to see the loss over interacting and non-interacting node pairs separately.
-These can be interpreted as analogue to the algorithms false negative rate and false positive rate.
+These can be interpreted as analogue to the algorithm's false negative rate and false positive rate:
 
@@
 	L^{(-)} &= \frac{1}{|\binom{V}{2} \setminus E_G|}\sum_{e \in \binom{V}{2} \setminus E_G} \prob(e \in E_G|X)^2
-\end{align}
+\end{align}.
 Other performance metrics such as the area under ROC curves or weighted sums over the quadratic loss functions above are also possible to apply.

### 2011-08-20 src/main/latex/paper/paper.tex
moved description of priors to earlier position in text

@@
 
+To determine the joint posterior distribution $\pi(\theta,G|X)$,  one may exploit the following equation:
+
+\begin{equation} 
+	 \pi(\theta,G|X) = \frac{\pi(\theta,G,X)}{\prob(X)} 
+\end{equation}.
+
+Thus, the posterior distribution $\pi(\theta,G|X)$ is proportional to the joint distribution $\pi(\theta,G,X)$. The joint distribution may in turn be determined as follows.
+
+\begin{equation} 
+	 \pi(\theta,G,X) = \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G)
+\end{equation}
+
+In summary, the following statement can be made.
+
+\begin{equation} 
+	 \pi(\theta,G|X) \propto \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G) 
+\end{equation}
+
+As a consequence, three values need to be determined: the prior distribution $\pi(G)$, the prior distribution $\pi(\theta)$ and the likelihood $\prob(X|\theta,G)$.
+
+We define the prior distribution of $\pi(G)$ based on the prior probability of each single edge to exist:
+\begin{align}
+  \pi(G) &= \prod_{e \in \binom{V}{2}} \pi(G_e)\\
+         &= (1-q)^{\left |\binom{V}{2}\setminus E_G \right |} \cdot q^{|E_G|}
+\end{align},
+where $q$ is the prior probability of an edge to really exist. \note{Not sure we really need this, since we don't use it afterwards anywhere.}
+
+To determine the prior distribution $\pi(\theta)$, we have to consider the nature of the error rates $\alpha_i$ and $\beta_i$ as ``success rates'' for misreading each potential edge.
+Modelling each observation event over a potential edge as a Bernoulli experiment with such a success rate, the number of false positives and false negatives in an experimental graph $X_i$ would follow a binomial distribution.
+The success rates for a binomial distribution can be sampled using the Beta distribution.
+The Beta distribution is dependent on two parameters, $a$ and $b$.
+
 However, instead of sampling from $\pi(\theta,G|X)$ directly, we employ Gibbs sampling \citep{gelfand_sampling-based_1990} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.
@@
 The second step in each cycle is the sampling of a new error rate vector $\theta$ based on $G$. 
-The error rates $\alpha_i$ and $\beta_i$ behave like the success rates for misreading each potential edge.
-Modelling each observation event over a potential edge as a Bernoulli experiment with such a success rate, the number of false positives and false negatives in an experimental graph $X_i$ would follow a binomial distribution.
-The success rates for a binomial distribution can be sampled using the Beta distribution.
-The Beta distribution takes two parameters, $a$ and $b$.
+%FIXME
 The parameters $a$ and $b$ can be interpreted as one plus the number of successes and failures, respectively, in a binomial distributed dataset.
@@
   \beta_i &\sim Be(\text{fn}_i+1, \text{tp}_i+1)
-\end{align}
+\end{align}.
 
@@
 %Finally, a score analogous to the likelihood of the current state can be computed as $\pi(\theta,G|X)$.
-%Lacking a way to determine this value directly, one may exploit the following equation.
-
-%\begin{equation} 
-%	 \pi(\theta,G|X) = \frac{\pi(\theta,G,X)}{\prob(X)} 
-%\end{equation}
-
-%Thus, the posterior distribution $\pi(\theta,G|X)$ is proportional to the joint distribution $\pi(\theta,G,X)$. The joint distribution may in turn be determined as follows.
-
-%\begin{equation} 
-%	 \pi(\theta,G,X) = \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G)
-%\end{equation}
-
-%In summary, the following statement can be made.
-
-%\begin{equation} 
-%	 \pi(\theta,G|X) \propto \pi(G)\cdot\pi(\theta)\cdot\prob(X|\theta,G) 
-%\end{equation}
-
-%As a consequence, three values need to be determined: the prior distribution $\pi(G)$, the prior distribution $\pi(\theta)$ and the likelihood $\prob(X|\theta,G)$.
 
@@
 
+\note{TODO: Fix capitalization and italics in BibTeX file}
+
 \end{document}

### 2011-08-20 src/main/latex/paper/paper.tex
added description of beta priors

@@
 The Beta distribution is dependent on two parameters, $a$ and $b$.
+For the prior distribution we assume $a = b = 1$.
 
-However, instead of sampling from $\pi(\theta,G|X)$ directly, we employ Gibbs sampling \citep{gelfand_sampling-based_1990} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.
+\begin{align}
+  \alpha_i &\sim Be(a_\alpha, b_\alpha) \; \forall i=1,\ldots,n\\
+  \beta_i &\sim Be(a_\beta, b_\beta) \; \forall i=1,\ldots,n
+\end{align}.
+
+Since sampling from $\pi(\theta,G|X)$ directly would be very difficult, we instead employ a Gibbs sampling approach \citep{gelfand_sampling-based_1990} and alternately sample from $\pi(\theta|G,X)$ and $\pi(G|\theta,X)$.
 
@@
 The second step in each cycle is the sampling of a new error rate vector $\theta$ based on $G$. 
-%FIXME
+As explained above, Beta distributions can be used to describe the behaviour of $\theta$.
 The parameters $a$ and $b$ can be interpreted as one plus the number of successes and failures, respectively, in a binomial distributed dataset.

### 2011-08-20 src/main/latex/paper/paper.tex
extending results section

@@
 As a consequence, it is hard to differentiate between node pairs that have been observed not to interact and node pairs that have not been subject to testing.
-\note{TODO: Low-throughput experiments have yet to be implemented}%TODO <--
 
@@
 \end{align}.
-Other performance metrics such as the area under ROC curves or weighted sums over the quadratic loss functions above are also possible to apply.
-However we find that the separate loss functions shown above give a more detailed picture of a method's behaviour.
-\note{Maybe move to discussion. Discuss other possible scoring functions, e.g. weighted sums over FPs and FNs}
 
@@
 Figure \ref{htp} shows the methods' performances over different numbers of input HTP experiments.
-It is clearly visible that the new fully Bayesian method outperforms both gold standard-based methods, particularly given a low number of input datasets.
+It is clearly visible that the new fully Bayesian method overall outperforms both gold standard-based methods.
+For low numbers of input datasets, the loss over existing interactions is significantly lower.
+Given more than five HTP experiments, all methods show comparably low loss of existing interactions.
+
+Regarding non-existing interactions the fully Bayesian method constantly performs better than the method presented by James and colleagues.
+\note{See if we can get rid of Lee's strange behaviour here, by using a different threshold.}
+
+The average loss regarding the $\theta$ estimates shows the fully Bayesian method's superior precision, which increases with the number of input experiments, while both gold standard-based methods stagnate at constant large amout of loss.
+
+\note{TODO: re-run on biased data}
+
+\note{TODO: Show some stats on the integration of real data}
 
@@
 
-When given a mixed set of HTP and LTP experiments, ...\note{TODO: Implement LTP simulation}
+%When given a mixed set of HTP and LTP experiments, ...\note{TODO: Implement LTP simulation}
 
 \section{Discussion}
+
+%In addition to the performance metrics used here, others such as the area under ROC curves or weighted sums over the quadratic loss functions above are also possible to apply.
+However we find that the separate loss functions shown above give a more detailed picture of a method's behaviour.
+%\note{Discuss other possible scoring functions, e.g. weighted sums over FPs and FNs}
 

### 2011-08-23 src/main/r/plot.R
probability histograms

@@ -18,3 +18,8 @@ barplot(as.matrix(t(er)),beside=TRUE,xlab="# experiments",ylab=expression(E(L["E
 par(op)
 dev.off()
 
+pdf("plot.pdf",width=4,height=4)
+probs = read.csv("probabilities.csv")
+hist(probs[,1],main="",xlab="Edge probability")
+dev.off()
+

### 2011-08-23 src/main/r/plot.R
results split up by PPI and GI

@@ -1,6 +1,6 @@
 #!/usr/bin/Rscript
 
-pdf("plot.pdf",width=4,height=8)
+pdf("fig1.pdf",width=4,height=8)
 op=par(mfrow=c(3,1))
 
 fn <- read.table("fn.tsv",header=TRUE,row.names=1,sep="\t")
@@ -18,8 +18,12 @@ barplot(as.matrix(t(er)),beside=TRUE,xlab="# experiments",ylab=expression(E(L["E
 par(op)
 dev.off()
 
-pdf("plot.pdf",width=4,height=4)
-probs = read.csv("probabilities.csv")
-hist(probs[,1],main="",xlab="Edge probability")
+pdf("fig2.pdf",width=6,height=12)
+op=par(mfrow=c(2,1))
+probsppi = read.csv("probabilities_ppi.csv")
+hist(probsppi[,1],main="",xlab="PPI probability")
+probsgi = read.csv("probabilities_gi.csv")
+hist(probsgi[,1],main="",xlab="GI probability")
+par(op)
 dev.off()
 

### 2011-08-24 src/main/latex/paper/paper.tex
results split up by PPI and GI
Short addition to results

@@
 
+\subsection{Evaluation on simulated data}
+
 Figure \ref{htp} shows the methods' performances over different numbers of input HTP experiments.
@@
 
-\note{TODO: Show some stats on the integration of real data}
-
 \begin{figure}[!tpb]
@@
 \label{htp}
+\end{figure}
+
+\subsection{Application to physical and genetic interactions}
+
+Figure \ref{ppi+gi} shows histograms of the existence probabilities for physical and genetic interactions from the semantically integrated dataset as calculated by the method presented here.
+It is apparent that \note{NUMBER}\% of the experimentally reported protein-protein interactions have been assigned probabilites less than 0.1.
+This is consistent with the reports of \note{PAPER}.
+
+\begin{figure}[!tpb]
+\centerline{\includegraphics[width=.5\textwidth]{fig2.pdf}}
+\caption{Histogram of existence probabilities for the interactions in the integrated dataset. Top: Protein-protein interactions, bottom: genetic interactions.}
+\label{ppi+gi}
 \end{figure}

### 2011-08-27 src/main/java/uk/ac/ncl/cs/harness/experiments/BiasedExperiment.java
implemented biased experiment simulation

@@ -1,0 +1,186 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ *
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ *
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import cern.jet.random.Beta;
+import cern.jet.random.Uniform;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.Map;
+import java.util.HashMap;
+import java.util.logging.Logger;
+import uk.ac.ncl.cs.harness.extsupport.Extension;
+import uk.ac.ncl.cs.harness.extsupport.ExtensionProperty;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.Node;
+import uk.ac.ncl.cs.harness.stats.RandomEngineRegistry;
+import uk.ac.ncl.cs.harness.util.SetOfTwo;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc.
+ */
+@Extension(id="biasedexperiment")
+public final class BiasedExperiment implements Experiment {
+
+    /**
+     * A random number generator.
+     */
+    private Uniform uniform = new Uniform(0.0, 1.0, RandomEngineRegistry.getEngine());
+
+    /**
+     * The experiment's sensitivity.
+     */
+    private double sensitivity;
+
+    /**
+     * The experiment's false discovery rate.
+     */
+    private double fdr;
+    
+    /**
+     * false positve rates for each non-edge.
+     */
+    private Map<SetOfTwo<Node>,Double> fprTable;
+    
+    /**
+     * false negative rate for each edge.
+     */
+    private Map<SetOfTwo<Node>,Double> fnrTable;
+
[…]

### 2011-08-27 src/test/java/uk/ac/ncl/cs/harness/experiments/BiasedExperimentTest.java
junit test for biased experiment

@@ -1,0 +1,120 @@
+/*
+ *  Copyright (C) 2011 Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ * 
+ *  This program is free software: you can redistribute it and/or modify
+ *  it under the terms of the GNU Lesser General Public License as published by
+ *  the Free Software Foundation, either version 3 of the License, or
+ *  (at your option) any later version.
+ * 
+ *  This program is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ * 
+ *  You should have received a copy of the GNU Lesser General Public License
+ *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+package uk.ac.ncl.cs.harness.experiments;
+
+import java.util.HashSet;
+import java.util.Set;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import junit.framework.TestCase;
+import uk.ac.ncl.cs.harness.LoggingInit;
+import uk.ac.ncl.cs.harness.graph.Edge;
+import uk.ac.ncl.cs.harness.graph.Graph;
+import uk.ac.ncl.cs.harness.graph.MapGraph;
+import uk.ac.ncl.cs.harness.graph.populators.ScaleFreePopulator;
+
+/**
+ *
+ * @author Jochen Weile, M.Sc. <j.weile@ncl.ac.uk>
+ */
+public class BiasedExperimentTest extends TestCase {
+
+    private Graph graph;
+
+    public BiasedExperimentTest(String testName) {
+        super(testName);
+    }
+
+    @Override
+    protected void setUp() throws Exception {
+        LoggingInit.init();
+
+        super.setUp();
+
+        graph = new MapGraph();
+        graph.setName("main");
+        ScaleFreePopulator populator = new ScaleFreePopulator();
+        populator.setSeed(3);
+        populator.setNumberOfNodes(500);
+        populator.populate(graph);
+
+        Logger.getLogger(this.getClass().getName())
+                .log(Level.INFO, "Template edges: {0}",
+                     graph.getNumEdges());
+    }
+
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+        graph = null;
+    }
+
+    /**
+     * Test of perform method, of class SimpleExperiment.
+     */
+    public void testPerform() {
+
+        
[…]

### 2011-08-28 src/test/java/uk/ac/ncl/cs/harness/experiments/BiasedExperimentTest.java
tested the bias simulation

@@ -90,15 +90,18 @@ public class BiasedExperimentTest extends TestCase {
         Logger.getLogger(this.getClass().getName())
                 .log(Level.INFO, "Sensitivity discrepancy: {0} -> {1}",
                      new Object[]{sens, actualSens});
+        System.out.println("Sensitivity discrepancy: "+sens+" -> "+actualSens);
 
         Logger.getLogger(this.getClass().getName())
-                .log(Level.INFO, "FPR discrepancy: {0} -> {1}",
+                .log(Level.INFO, "FDR discrepancy: {0} -> {1}",
                      new Object[]{fdr, actualFdr});
+        System.out.println("FDR discrepancy: "+fdr+" -> "+actualFdr);
 
 
         Logger.getLogger(this.getClass().getName())
                 .log(Level.INFO, "Experimental edges: {0}",
                      experimentalGraph.getNumEdges());
+        System.out.println("Experimental edges: "+experimentalGraph.getNumEdges());
     }
 
      /**

### 2011-08-29 src/main/latex/paper/paper.tex
paper: small changesto results, gathering ideas for discussion

@@
 
+\note{TODO: Discuss value of prior odds}
+
 In order to calculate the Bayes factors above it is necessary to determine the rates of false positives and false negatives in each dataset.
@@
 This type of error is easily simulated by randomly inserting edges that do not exist in the real data and removing edges that do exist in the real data until the desired error rates are reached.
-The next problem is systematic error or experimental bias.
+The next problem is systematic error, also known as experimental bias.
 This phenomenon in particular leads to the violation of stochastic independence between datasets.
-\note{TODO: Bias simulation has yet to be implemented.}%TODO <-- 
-Finally, another type of error that easily escapes one's attention is the lack of negative records.
+We simulate this by sampling separate false negative probabilities for each interaction and false positive probabilities for each non-interacting node-pair from Beta distributions.
+We then use these prepared probilities to generate false positives and false negatives in the experiments, thus introducing the same bias/systematic error.
+Pseudocode for this method is provided in the supplementary material.
+Finally, another type of problem that easily escapes one's attention is the lack of negative records.
 Most experiments do not report which pairs of nodes do not possess an edge.
@@
 
-\note{TODO: re-run on biased data}
-
 \begin{figure}[!tpb]
@@
 Figure \ref{ppi+gi} shows histograms of the existence probabilities for physical and genetic interactions from the semantically integrated dataset as calculated by the method presented here.
-It is apparent that \note{NUMBER}\% of the experimentally reported protein-protein interactions have been assigned probabilites less than 0.1.
-This is consistent with the reports of \note{PAPER}.
+A share of 39.03\% of the experimentally reported protein-protein interactions have been assigned probabilites less than 0.1.
+This percentage is lower than what could be expected according to the projections of Hart and colleagues \citep{hart_how_2006}.
 
@@
 \section{Discussion}
+\note{
+\begin{itemize}
+  \item Excellent performance despite giving competitors head start with unbiased gold standards and accurate fudge factors, which wouldn't be available normally.
+  \item Other possible performance metrics (e.g. ROC curves or weighted sums)
+  \item Discuss discrepancy between Hart et al and the PPI network
+  \item Future work: Evaluate on simulated low-throughput
+  \item Future work: Further evaluation of the algorithm using the integrated dataset
+  \begin{itemize}
+    \item Measure functional annotation accuracy with leave-one-out
+  \end{itemize}
+\end{itemize}
+}
 
@@
 %\note{Discuss other possible scoring functions, e.g. weighted sums over FPs and FNs}
-
-\note{TODO: Write discussion}
 

### 2011-08-29 src/main/r/plot.R

@@ -18,12 +18,12 @@ barplot(as.matrix(t(er)),beside=TRUE,xlab="# experiments",ylab=expression(E(L["E
 par(op)
 dev.off()
 
-pdf("fig2.pdf",width=6,height=12)
+pdf("fig2.pdf",width=6,height=9)
 op=par(mfrow=c(2,1))
 probsppi = read.csv("probabilities_ppi.csv")
-hist(probsppi[,1],main="",xlab="PPI probability")
+hist(probsppi[,1],main="",xlab="PPI probability",col="gray")
 probsgi = read.csv("probabilities_gi.csv")
-hist(probsgi[,1],main="",xlab="GI probability")
+hist(probsgi[,1],main="",xlab="GI probability",col="gray")
 par(op)
 dev.off()
 

### 2011-09-05 src/main/r/plot.R
synthetic lethal data

@@ -22,8 +22,8 @@ pdf("fig2.pdf",width=6,height=12)
 op=par(mfrow=c(2,1))
 probsppi = read.csv("probabilities_ppi.csv")
 hist(probsppi[,1],main="",xlab="PPI probability")
-probsgi = read.csv("probabilities_gi.csv")
-hist(probsgi[,1],main="",xlab="GI probability")
+probsgi = read.csv("probabilities_sl.csv")
+hist(probsgi[,1],main="",xlab="Synthetic GI probability")
 par(op)
 dev.off()
 

### 2011-09-05 src/main/r/plot.R
updated paper figures with synthetic GI data

@@ -18,7 +18,7 @@ barplot(as.matrix(t(er)),beside=TRUE,xlab="# experiments",ylab=expression(E(L["E
 par(op)
 dev.off()
 
-pdf("fig2.pdf",width=6,height=9)
+pdf("fig2.pdf",width=6,height=8)
 op=par(mfrow=c(2,1))
 probsppi = read.csv("probabilities_ppi.csv")
 hist(probsppi[,1],main="",xlab="PPI probability",col="gray")

### 2011-09-05 src/main/latex/paper/paper.tex
updated paper figures with synthetic GI data
paper: some more results, some more literature

@@
 However, gold standards only exist for some types of associations.
-Therefore, it is desirable to find a method that can infer confidence measures on biological networks in the absence of a gold standard, that is, based only on the existing experimental data.
+Therefore, it is desirable to find a method that can infer confidence measures on biological networks, preferrably in the absence of a gold standard, that is, based only on the existing experimental data.
 So far, only partial solutions to this problem exist.
-These solutions infer confidence measures for a specific type of association only, for example, protein-protein interactions \citep{bader_gaining_2004}.\note{Find more}
+These solutions infer confidence measures for a specific type of association only, for example, protein-protein interactions \citep{bader_gaining_2004,braun_experimentally_2009,venkatesan_empirical_2009}.
 
@@
 A description of the algorithm can be found in the supplementary material.
-\note{Mention how this does not match the algorithm's definition of the prior distribution of $G$, thus making it more challenging for the algorithm to perform.}%TODO
+Not only does the choice of scale free background graphs more closely match the topology of real biological graphs; it also poses an additional challenge for the new algorithm, since such graphs break with the assumption of a uniform prior distribution for $G$.
 
@@
 
-\subsection{Semantic integration}
+\subsection{Application to biological networks}
 
-In addition to the evaluation on artificial data it is important to analyse the method's performance on real biological data.
-To this end we have used the semantic data integration system Ondex \citep{koehler_graph-based_2006} to gather as much data as possible on the \species{Saccharomyces cerevisiae} protein-protein interaction (PPI) network.
-A new PSI-MI 2.5 XML \citep{kerrien_broadening_2007} parser was written to import data from the BioGRID \citep{breitkreutz_biogrid_2008}, MINT \citep{chatr-aryamontri_mint:_2007}, IntAct \citep{hermjakob_intact:_2004} and MIPS-MPACT \citep{guldener_mpact:_2006}.
-The resulting Ondex dataset represented proteins, their interactions, the experiments in which the interactions have been observed and the publications in which the experiments were described.
-A semantic merger method was developed to clean the dataset of its redundancies.
+In addition to the evaluation on artificial data we applied the new method to semantically integrated biological data.
+We used the data integration system Ondex \citep{koehler_graph-based_2006} to gather as much data as possible on the \species{Saccharomyces cerevisiae} protein-protein interaction (PPI) and genetic interaction (GI) network.
+W have imported all\species{S.
+cerevisiae} data from the BioGRID \citep{breitkreutz_biogrid_2008}, MINT \citep{chatr-aryamontri_mint:_2007}, IntAct \citep{hermjakob_intact:_2004} and MIPS-MPACT \citep{guldener_mpact:_2006} via the PSI-MI 2.5 XML \citep{kerrien_broadening_2007} format.
+The resulting Ondex dataset represented proteins and genetic features, as well as their interactions, the experiments in which the interactions have been observed and the publications in which the experiments were described.
+To identify and interlink equivalent entries we implemented a semantic merger method.
+Finally, we summarized small low-throughput experiments (<20 interactions) into larger groups according to their experimental type, since low sample sizes can be expected to lead to very unprecise error rate estimates.
 Further details on the semantic integration procedure can be found in the supplementary material.
+
+Having established a semantic knowledge network of physical and genetic interactions within \species{S.
+cerevisiae} We executed the new algorithm over all 51,006 contained physical interactions (PSI-MI ontology ID MI:0915) and all 15,006 contained synthetic genetic interactions (PSI-MI ontology ID MI:0794), as an exemplary type of genetic interaction.
 
@@
 
-\subsection{Application to physical and genetic interactions}
+\subsection{Interaction probabilities for \species{S. cerevisiae}}
 
-Figure \ref{ppi+gi} shows histograms of the existence probabilities for physical and genetic interactions from the semantically integrated dataset as calculated by the method presented here.
-A share of 39.03\% of the experimentally reported protein-protein interactions have been assigned probabilites less than 0.1.
-This percentage is lower than what could be expected according to the projections of Hart and colleagues \citep{hart_how_2006}.
+Figure \ref{ppi+gi} shows histograms of the existence probabilities for physical interactions and synthetic genetic interactions from the semantically integrated dataset as calculated by the method presented here.
+Regarding physical interactions, 20,180 (39.03\%) of the experimentally reported interactions have been assigned probabilites less than 0.1.
+A total of 22090 (43.3\%) have been assigned probabilites greater than 0.9.
+The remaining interactions are widely distributed around a small peak near probability 0.3.
+
+\note{TODO: Output error rates and compare with values from \citep{dhaeseleer_estimating_2004} and \citep{hart_how_2006}}
+
+When looking at the probability distribution over synthetic genetic interactions it is apparent that with 14,386 (95.8\%) a vast majority of interactions have been assigned values of less than 0.5.
+\note{Potential explanation: Extremely poor coverage.
+TODO: run coverage analysis in Ondex to confirm.} Of these, 8086 (53.9\%) have a probability less than 0.1.
+Only 230 (1.5\%) have a probability greater than 0.9.
+
+%\citep{hart_how_2006} summarizes literature about PPI interactome size projections as between 37,800-75,500.
 
@@
 \centerline{\includegraphics[width=.5\textwidth]{fig2.pdf}}
-\caption{Histogram of existence probabilities for the interactions in the integrated dataset. Top: Protein-protein interactions, bottom: genetic interactions.}
+\caption{Histogram of existence probabilities for the interactions in the integrated dataset. Top: Protein-protein interactions, bottom: synthetic genetic interactions.}
 \label{ppi+gi}
 \end{figure}
-
-%When given a mixed set of HTP and LTP experiments, ...\note{TODO: Implement LTP simulation}
 
@@
   \item Other possible performance metrics (e.g. ROC curves or weighted sums)
-  \item Discuss discrepancy between Hart et al and the PPI network
   \item Future work: Evaluate on simulated low-throughput
@@
   \end{itemize}
+  \item Future work: include confidence values given by original experiments. (Discuss problem of missing standards)
 \end{itemize}
@@
 
-\note{TODO: Fix capitalization and italics in BibTeX file}
+\note{TODO: Fix capitalization and italics in BibTeX file (Adding items from Zotero will undo this, so wait until the end)}
 

### 2011-09-11 src/main/latex/paper/paper.tex
More related work in introduction

@@
 We present a novel algorithm for computing the probability of experimental observations to be correct that operates without gold standard reference data.
-We show that our algorithm outperforms existing gold standard based methods.
-Finally, we apply the new algorithm on a large collection of genetic interaction and protein-protein interaction experiments.
+We show that our algorithm outperforms existing gold standard-based methods.
+Finally, we apply the new algorithm to a large collection of genetic interaction and protein-protein interaction experiments.
 
@@
 
-A popular integration paradigm is the construction of functional networks \citep{von_mering_string:_2003,warde-farley_genemania_2010}\note{Also cite Katherine}.
+A popular integration paradigm is the construction of functional networks \citep{lee_probabilistic_2004,von_mering_string:_2003,james_integration_2009}.
 Functional networks represent different types of relationships between biological entities in an abstract manner.
@@
 
-An alternative paradigm is semantic data integration \citep{cheung_yeasthub:_2005,smith_obo_2007,koehler_graph-based_2006}.
+An alternative paradigm is semantic data integration \citep{cheung_yeasthub:_2005,smith_obo_2007,koehler_graph-based_2006}\note{Cite Pathway commons here}.
 These approaches aim at representing the biological information (and as much of its meaning as possible) in a computationally accessible fashion.
@@
 
-An important question in data integration is how to assess the degree of confidence in each statement, that is, how likely the statement is to be correct.
-Several popular solutions to this problem exist in the context of functional networks \citep{lee_probabilistic_2004,troyanskaya_bayesian_2003}.
+An important question regarding such networks is how to assess the degree of confidence in each statement, that is, how likely the statement is to be correct.
+Several popular solutions to this problem exist for functional networks \citep{lee_probabilistic_2004}\note{REF Lycett when it's accepted}.
 These methods assess the quality of each input dataset against one or more additional datasets of higher quality, usually manually-curated collections.
@@
 
-Inferring confidence assessments for semantically integrated datasets is challenging, because each single type of association must be scored separately.
-However, gold standards only exist for some types of associations.
-Therefore, it is desirable to find a method that can infer confidence measures on biological networks, preferrably in the absence of a gold standard, that is, based only on the existing experimental data.
-So far, only partial solutions to this problem exist.
-These solutions infer confidence measures for a specific type of association only, for example, protein-protein interactions \citep{bader_gaining_2004,braun_experimentally_2009,venkatesan_empirical_2009}.
+The method described by Lee and colleagues evaluates each evidential dataset against such a gold standard and obtain a log likelihood score (LLS).
+Subsequently, for each interaction in question, a weighted sum is formed over the LLS scores of those datasets that report the interaction.
+The weights are chosen in a manner that represents the degree of dependency between the datasets \citep{lee_probabilistic_2004}.
 
-%Whether or not this description is justified is a matter of dispute \REF.
+James and colleagues describe a method that extends the original method of Lee and colleagues \note{Cite Sam here as soon as the paper is out}.
+Not only one, but several different gold standards are used to generate LLS scores for the datasets.
+Furthermore, instead of creating a score for each interaction via the weighted sum described above, this method computes an existence probability from the original LLS scores and then averages over the different probabilites according to the different gold standards.
+The authors show that any bias inherent in the used gold standards can thus be overcome.
+\note{REF Lycett paper again here}
 
-%It would be desirable to investigate these ideas at a more detailed level, that is, moving on from the question: ``How likely are two given proteins to be functionally related?'', to more specific questions such as: ``How likely is it that two given proteins physically interact?''.
-However, at this level it becomes infeasible to rely on gold standards, \note{since the level of completeness and quality required has not yet been achieved.} simply because none exist \REF.
-Therefore, it is necessary to develop a method that can infer confidence measures for semantic statements about molecular biological entities, based only on the existing data.
+These methods work very well for functional networks.
+However, inferring confidence assessments for semantic networks, rather than functional networks, is more challenging, because each single type of association must be scored separately.
+Reliable gold standards only exist for some of these types.
+The methods discussed above are thus only of limited use for assessing semantic networks.
+While solutions for specific types of data do exist, for example, protein-protein interactions \citep{troyanskaya_bayesian_2003,bader_gaining_2004,braun_experimentally_2009,venkatesan_empirical_2009}, These are again dependent on additional data.
+It would desirable to find a method that can infer confidence measures on biological networks in the absence of a gold standard, that is, based only on the existing experimental data.
 

### 2011-09-12 src/main/latex/paper/paper.tex
discussion

@@
 
-James and colleagues describe a method that extends the original method of Lee and colleagues \note{Cite Sam here as soon as the paper is out}.
+James and colleagues describe a method that extends the original method of Lee and colleagues.
 Not only one, but several different gold standards are used to generate LLS scores for the datasets.
@@
 The Beta distribution is dependent on two parameters, $a$ and $b$.
-For the prior distribution we assume $a = b = 1$.
+For the sake of simplicity, we here assume $a = b = 1$ for the prior distribution.
 
@@
 
-To initiate the algorithm as a whole, we need to generate initial error rate vector $\theta_{\text{init}}$.
+To initiate the algorithm as a whole, we need to generate an initial error rate vector $\theta_{\text{init}}$.
 It is sufficient to sample the initial values for each $\alpha_i$ and $\beta_i$ from their prior distribution.
@@
 In the simulation these graphs assume the role of experimental datasets.
-
 The simulated experimental datasets are subsequently passed to the integration method under investigation.
 The integrated graph resulting from the integration method is then compared with the original simulated true graph to evaluate the integration method's performance.
+
 Such a testing workflow can be programmed to be executed a large number of times in order to measure a method's average behaviour.
@@
 This type of error is easily simulated by randomly inserting edges that do not exist in the real data and removing edges that do exist in the real data until the desired error rates are reached.
+
 The next problem is systematic error, also known as experimental bias.
@@
 Pseudocode for this method is provided in the supplementary material.
-Finally, another type of problem that easily escapes one's attention is the lack of negative records.
+\note{TODO: Make sure it is}
+
+%Finally, another type of problem that easily escapes one's attention is the lack of negative records.
 Most experiments do not report which pairs of nodes do not possess an edge.
@@
 There are a number of different quality measures that can be applied.
-One important aspect is to measure the accuracy of the method in estimating the error rates in the individual experiments.
+One important aspect is to measure the accuracy of error rates estimates in the individual experiments.
 We can define the quadratic loss for the error estimates as follows:
@@
 We have evaluated the new algorithm in comparison with two gold standard-based methods presented by Lee and colleagues \citep{lee_probabilistic_2004} and James and colleagues \REF.
-For our evaluation we have examined the following scenario: Each integration method is given the task of processing a number (3,5,7,9,11) of high-throughput (HTP) experiments with average false negative rate $0.15$ and average false discovery rate of $0.15$ (corresponds to false positive rate of 0.0006).
-In the case of the gold standard-based methods, one of the input experiments serves as the gold standard.
+For our evaluation we have examined the following scenario: Each integration method is given the task of processing a number (3,5,7,9, and 11) of experimental scale-free networks with 500 nodes, an average false negative rate $0.15$ and average false discovery rate of $0.15$ (corresponds to false positive rate of 0.0006).
+The two gold standard-based methods have been given a bias-free input experiment as the gold standard and their additional input parameters (such as the dependency factor) have been set to optimal values to ensure peak performance.
+We have run 5000 replicates of the above test and averaged over the results.
 
@@
 We used the data integration system Ondex \citep{koehler_graph-based_2006} to gather as much data as possible on the \species{Saccharomyces cerevisiae} protein-protein interaction (PPI) and genetic interaction (GI) network.
-W have imported all\species{S.
+We imported all\species{S.
 cerevisiae} data from the BioGRID \citep{breitkreutz_biogrid_2008}, MINT \citep{chatr-aryamontri_mint:_2007}, IntAct \citep{hermjakob_intact:_2004} and MIPS-MPACT \citep{guldener_mpact:_2006} via the PSI-MI 2.5 XML \citep{kerrien_broadening_2007} format.
@@
 Further details on the semantic integration procedure can be found in the supplementary material.
+\note{TODO: Make sure it is.}
 
@@
 
-Figure \ref{htp} shows the methods' performances over different numbers of input HTP experiments.
+Figure \ref{htp} shows the three methods' performances over different numbers of input experiments.
 It is clearly visible that the new fully Bayesian method overall outperforms both gold standard-based methods.
 For low numbers of input datasets, the loss over existing interactions is significantly lower.
-Given more than five HTP experiments, all methods show comparably low loss of existing interactions.
+Given more than five experiments, all methods show comparably low loss of existing interactions.
 
@@
 
-The average loss regarding the $\theta$ estimates shows the fully Bayesian method's superior precision, which increases with the number of input experiments, while both gold standard-based methods stagnate at constant large amout of loss.
+The average loss regarding the $\theta$ estimates shows the fully Bayesian method's superior precision, which increases with the number of input experiments, while both gold standard-based methods stagnate at constant large amounts of loss.
 
@@
 \section{Discussion}
-\note{
-\begin{itemize}
-  \item Excellent performance despite giving competitors head start with unbiased gold standards and accurate fudge factors, which wouldn't be available normally.
-  \item Other possible performance metrics (e.g. ROC curves or weighted sums)
-  \item Future work: Evaluate on simulated low-throughput
-  \item Future work: Further evaluation of the algorithm using the integrated dataset
-  \begin{itemize}
-    \item Measure functional annotation accuracy with leave-one-out
-  \end{itemize}
-  \item Future work: include confidence values given by original experiments. (Discuss problem of missing standards)
-\end{itemize}
-}
+\subsection{Summary and conclusion}
 
-%In addition to the performance metrics used here, others such as the area under ROC curves or weighted sums over the quadratic loss functions above are also possible to apply.
-However we find that the separate loss functions shown above give a more detailed picture of a method's behaviour.
-%\note{Discuss other possible scoring functions, e.g. weighted sums over FPs and FNs}
+We have presented a novel, fully Bayesian method for assessing the credibility of semantically integrated experimental network data in the absence of gold standards.
+We have evaluated the method's performance in comparison to two existing gold standard-based methods using realistically simulated data.
+The new method has shown excellent performance despite the testing environment being designed to specifically challenge it: both competing reference methods have been given optimal input parameters as well as unbiased gold standards, which would not normally be available.
+Additionally the environment has been set up to simulate scale-free graphs, which emulate the topology of real biological graphs and thus do not meet the algoriths assumptions regarding the prior distribution of $G$.
+Finally, the experiment simulation has been designed to introduce bias and thus violate the algorithms independence assumption.
+
+We have evaluated the performance of the new algorithm as well as the two reference methods using three different metrics, which express the accuracy of error rate estimation as well as the accuracy of probability assignment to existing and non-existing edges.
+Other perfomance metrics could have been used as well (e.g.
+ROC-AUC or weighted sums of loss functions), but would have provided less detail.
+
+While we have shown that the new algorithm performs well on large input experiments, it remains yet to be shown how the method performs on smaller datasets originating from low-throughput experiments.
+A further limitation to the approach is that it does not take into account any potential pre-existing confidence assessments from the original experiments.
+
+We conclude that the new fully Bayesian method is a valuable addition to the set of tools available for confidence assessment of semantically integrated datasets.
+
+\subsection{Outlook}
+
+Future work will include a further evalation of the method on real biological data.
+It would be possible to compare the algorithm's performance to other methods when using the calculated confidence values for protein function prediction in a leave-one-out evaluation.
+
+Furthermore we would like to analyze the method's performance on small low-throughput type input data sets.
+Finally, it would be interesting to explore more of the huge simulation parameter space: A great number of combinations between different numbers of small and large input experiments with varying false positive and false negative rates remains to be surveyed.
+
+Another useful feature would be the incorporation of potential pre-existing confidence values.
+This is largely complicated by the lack of a common standard.
+For example, protein-protein interaction data as available in PSI-MI format from the databases mentioned above contains various different confidence values in different metrics, such as probabilities, letter grades or bit-scores.
 

### 2011-09-12 src/main/latex/paper/paper.tex
added pseudocode to paper

@@
 \usepackage{color} 
+\usepackage{float}
+\usepackage{forloop}
 
@@
 \newcommand{\REF}{{\color{red}[REF]}}
+%% Pseudocode macros %%
+\floatstyle{ruled}
+\newfloat{pseudocode}{thp}{lop}
+\floatname{pseudocode}{Pseudocode}
 
@@
 
-An alternative paradigm is semantic data integration \citep{cheung_yeasthub:_2005,smith_obo_2007,koehler_graph-based_2006}\note{Cite Pathway commons here}.
+An alternative paradigm is semantic data integration \citep{cheung_yeasthub:_2005,smith_obo_2007,koehler_graph-based_2006,cerami_pathway_2010}.
 These approaches aim at representing the biological information (and as much of its meaning as possible) in a computationally accessible fashion.
@@
 
-A full proof of the above equation is provided in the supplementary material. This recursive equation can be expressed iteratively as follows:
+A full proof of the above equation is provided in the supplementary methods. This recursive equation can be expressed iteratively as follows:
 
@@
 That is, the odds of the edge to exist given all the experimental measurements is the product of the Bayes factors for these measurements with the prior odds of edge existence.
-These odds can be converted into the equivalent probability.
+(The determination the prior odds are described in the supplementary methods.) These odds can be converted into the equivalent probability using the logistic function.
 
@@
 This assumption is not valid for real data, and thus introduces a potential source of error into the methodology.
-Lee and colleagues address this problem by introducing dependency co-efficients for their datasets, a solution which relies heavily on guesswork \citep{lee_probabilistic_2004}.
+Lee and colleagues address this problem by introducing dependency co-efficients for their datasets \citep{lee_probabilistic_2004}, a solution which we argue relies on user guesswork.
 Instead we focus on making our method robust towards this kind of error.
-
-\note{TODO: Discuss value of prior odds}
 
@@
 %$Be(1,1)$, which is equivalent to the uniform distribution $\mathcal{U}_{[0,1]}$. 
-Pseudocode for the complete algorithm is given in the supplementary material.
+%Pseudocode for the complete algorithm is given in the supplementary methods.
 
-%Finally, a score analogous to the likelihood of the current state can be computed as $\pi(\theta,G|X)$.
+\begin{pseudocode}
+  \scriptsize \input{pseudocode.tex}
+  \caption{The Gibbs sampling algorithm for sampling from $\pi(\theta,G|X)$.}
+\end{pseudocode}
 
@@
 The simulated true graph is created as a scale-free graph using a preferential attachment algorithm, since most molecular-biological graphs have been shown to be approximately scale-free \citep{jeong_lethality_2001, eisenberg_preferential_2003}.
-A description of the algorithm can be found in the supplementary material.
+A description of the algorithm can be found in the supplementary methods.
 Not only does the choice of scale free background graphs more closely match the topology of real biological graphs; it also poses an additional challenge for the new algorithm, since such graphs break with the assumption of a uniform prior distribution for $G$.
@@
 We then use these prepared probilities to generate false positives and false negatives in the experiments, thus introducing the same bias/systematic error.
-Pseudocode for this method is provided in the supplementary material.
+Pseudocode is provided in the supplementary methods.
 \note{TODO: Make sure it is}

### 2011-09-18 src/main/r/plot.R
adding naive method to results

@@ -5,7 +5,7 @@ op=par(mfrow=c(3,1))
 
 fn <- read.table("fn.tsv",header=TRUE,row.names=1,sep="\t")
 barplot(as.matrix(t(fn)),beside=TRUE,xlab="# experiments",ylab=expression(E(L^{"("+")"})) )
-legend("topright",c(expression(paste("Lee ",italic("et al."))),expression(paste("James ",italic("et al."))),expression(paste("Weile ",italic("et al.")))),bty="n",fill=grey.colors(3))
+legend("topright",c("Naive", expression(paste("Lee ",italic("et al."))),expression(paste("James ",italic("et al."))),expression(paste("Weile ",italic("et al.")))),bty="n",fill=grey.colors(4))
 
 fp <- read.table("fp.tsv",header=TRUE,row.names=1,sep="\t")
 barplot(as.matrix(t(fp)),beside=TRUE,xlab="# experiments",ylab=expression(E(L^{"("-")"})) )

### 2011-09-18 src/main/latex/paper/paper.tex
Improved explanation of input parameters, added naive method to results

@@
 For our evaluation we have examined the following scenario: Each integration method is given the task of processing a number (3,5,7,9, and 11) of experimental scale-free networks with 500 nodes, an average false negative rate $0.15$ and average false discovery rate of $0.15$ (corresponds to false positive rate of 0.0006).
-The two gold standard-based methods have been given a bias-free input experiment as the gold standard and their additional input parameters (such as the dependency factor) have been set to optimal values to ensure peak performance.
+The two gold standard-based methods have each been given one of these input experiment with the same rates, but freed from bias, as the gold standard.
+The additional input parameters for the gold standard-based methods (such as the dependency factor) have been set to optimal values to ensure peak performance.
 We have run 5000 replicates of the above test and averaged over the results.
+As a basic benchmark, we have also executed the same workflow for a naive integration method.
+The naive method simply assigns the share of experiments that support a given edge as its existence probability.
 
@@
 It is clearly visible that the new fully Bayesian method overall outperforms both gold standard-based methods.
-For low numbers of input datasets, the loss over existing interactions is significantly lower.
+For low numbers of input datasets, the fully Bayesian method's loss over existing interactions is significantly lower and is the only method to perform better than the naive approach for three input experiments.
 Given more than five experiments, all methods show comparably low loss of existing interactions.
@@
 \centerline{\includegraphics[width=.5\textwidth]{fig1.pdf}}
-\caption{Evaluation of the algorithm in comparison with GS-based methods by Lee and colleagues and James and colleagues given different numbers of HTP experiments.
+\caption{Evaluation of the algorithm in comparison with the naive method as well as the GS-based methods by Lee and colleagues and James and colleagues given different numbers of experiments.
 Top: Average loss over interacting node pairs ($L^{(+)}$).
 Centre: Average loss over non-interacting node pairs ($L^{(-)}$).
-Bottom: Average loss over regarding error rate estimates ($L_\text{ER}$).}
+Bottom: Average loss regarding error rate estimates ($L_\text{ER}$).}
 \label{htp}
